\hypertarget{the-lesson-from-the-past}{%
\subsection{The Lesson from the Past}\label{the-lesson-from-the-past}}

\hypertarget{section}{%
\section{1}\label{section}}

\bibverse{1} In the eighth month in the second year of Darius, this
message of the Lord came to the prophet Zechariah, the son of Berechiah,
the son of Iddo:

\bibverse{2} The Lord was furious with your ancestors. \bibverse{3} So
tell the people this message of the Lord of hosts: ``Return to me,''
says the Lord of hosts, ``and I will return to you'' says the Lord of
hosts. \bibverse{4} ``Do not be like your ancestors to whom the former
prophets cried, `The Lord of hosts says to turn now from your evil ways,
and from your evil deeds'. They did not heed nor listen to me,'' says
the Lord. \bibverse{5} ``Your ancestors, where are they? And the
prophets, do they live forever? \bibverse{6} But did not my words and my
statutes with which I charged my servants the prophets overtake your
ancestors? And they repented and said, `Just as the Lord of hosts
planned to do to us, according to our ways, and according to our deeds,
so he has dealt with us.'\,'' \#\# The Lord's Jealous Love for Zion

\bibverse{7} In the twenty-fourth day of the eleventh month, which is
the month Shebat, in the second year of Darius, this message of the Lord
came to the prophet Zechariah, the son of Berechiah, the son of Iddo:

\bibverse{8} I saw in the night a man on a red horse, amongst the myrtle
trees that were in the valley-bottom. Behind him there were horses, red,
sorrel, and white. \bibverse{9} Then said I, ``My Lord, what are
these?'' And the messenger who was talking with me said, ``I will show
you what they are.'' \bibverse{10} And the man who was standing amongst
the myrtle trees answered, ``They are those whom the Lord has sent to
patrol the earth.'' \bibverse{11} And the riders answered the messenger
of the Lord who was standing amongst the myrtle trees, ``We have gone up
and down through the world and behold, all the earth is still and at
peace.''

\bibverse{12} Then the messenger of the Lord said, ``Lord of hosts, how
long will you have no pity for Jerusalem and the cities of Judah with
which you have been angry for seventy years?'' \bibverse{13} And the
Lord answered the messenger who was talking with me with kind and
comforting words. \bibverse{14} So the messenger who was talking with me
said to me, ``Proclaim now that the Lord of hosts says: I am deeply
concerned for Jerusalem. \bibverse{15} But I am deeply angry with the
arrogant, complacent nations. I was only a little angry with Israel but
they made the suffering worse. \bibverse{16} Therefore, says the Lord, I
am turning to show mercy to Jerusalem. My temple will be rebuilt in it,
says the Lord of hosts, and a measuring line will be stretched over
Jerusalem. \bibverse{17} Proclaim again: Thus says the Lord of hosts: My
cities will again overflow with prosperity and the Lord will again
comfort Zion and choose Jerusalem.'' \#\# Destruction of the Heathen
Powers

\bibverse{18} Now I looked up and saw four horns. \bibverse{19} I said
to the messenger who was talking with me, ``What are these?'' He
answered, ``These are the horns which scattered Judah, Israel and
Jerusalem.''

\bibverse{20} Then the Lord showed me four smiths. \bibverse{21} I said,
``What are these coming to do?'' He answered, ``The horns scattered
Judah, so that no one dared lift their head. But the smiths have come to
terrify them, to crush the horns of the nations which lifted up their
horn against the land of Judah to scatter its people.''

\hypertarget{promise-that-jerusalem-will-be-gloriously-restored}{%
\subsection{Promise that Jerusalem Will be Gloriously
Restored}\label{promise-that-jerusalem-will-be-gloriously-restored}}

\hypertarget{section-1}{%
\section{2}\label{section-1}}

\bibverse{1} Then I looked up and saw a man with a measuring line in his
hand. \bibverse{2} I said, ``Where are you going?'' He replied, ``To
measure Jerusalem, to see what is its breadth and length.'' \bibverse{3}
Then the messenger who talked with me stepped forward, and another
messenger went out to meet him, \bibverse{4} and said to him, ``Run and
tell to that young man that Jerusalem will be inhabited like a village
without walls, because so many people and cattle will live there.
\bibverse{5} For I, says the Lord, will be a wall of fire around her,
and I will be the glory in the midst of her. \bibverse{6} Up! Away! Flee
from the land of the north, says the Lord. For I have spread you abroad
as the four winds of the heavens, says the Lord. \bibverse{7} Up, escape
to Zion, you who live in Babylon. \bibverse{8} This is what the Lord of
hosts says to the nations which plundered you: Whoever touches you
touches the apple of my eye. \bibverse{9} I will raise my hand against
them. They will be plunder for their own slaves; and you will know that
the Lord of hosts has sent me. \bibverse{10} ``Sing and rejoice,
daughter of Zion! I am coming, I will live amongst you, says the Lord.
\bibverse{11} Many nations will join themselves to the Lord in that day,
and will be his people, and he will live amongst you. You will know that
the Lord of hosts has sent me to you. \bibverse{12} The Lord will
inherit Judah as his portion in the holy land. He will again choose
Jerusalem. \bibverse{13} Be silent, everyone, in the presence of the
Lord, because he is coming from his holy dwelling.''

\hypertarget{purification-of-the-priesthood-and-nation}{%
\subsection{Purification of the Priesthood and
Nation}\label{purification-of-the-priesthood-and-nation}}

\hypertarget{section-2}{%
\section{3}\label{section-2}}

\bibverse{1} Then he showed me Joshua, the high priest, standing before
the messenger of the Lord and the Adversary standing at his right hand
to accuse him. \bibverse{2} And the messenger of the Lord said to the
Adversary, ``May the Lord rebuke you, Adversary! The Lord, who has
chosen Jerusalem, rebuke you! Is not this man a brand plucked out of the
fire?'' \bibverse{3} Now Joshua was clothed with filthy garments and was
standing before the messenger. \bibverse{4} The messenger said to those
who stood before him, ``Take off his filthy garments.'' Then to Joshua
he said, ``I have taken your guilt from you and I will clothe you in
priestly robes.'' \bibverse{5} And he ordered: ``Set a clean turban upon
his head.'' So they set a clean turban on his head, and clothed him with
garments. The messenger of the Lord was standing by.

\bibverse{6} And the messenger of the Lord testified to Joshua:
\bibverse{7} The Lord of hosts says: ``If you walk in my ways, and if
you will keep my instructions, then you will rule my house and keep my
courts and I will give you a place of access amongst these that stand
by. \bibverse{8} Hear now, Joshua the high priest, you and your
associates who sit before me; for they are men who are a sign; for I am
going to bring forth my servant the Branch. \bibverse{9} On the stone
that I have set before Joshua, on one stone, are seven facets. I will
engrave it,'' says the Lord of hosts, ``and I will remove the guilt of
that land in one day. \bibverse{10} On that day,'' says the Lord of
hosts, ``you will each invite your neighbour under the vine and under
the fig tree.''

\hypertarget{the-supporters-of-the-temple-service}{%
\subsection{The Supporters of the Temple
Service}\label{the-supporters-of-the-temple-service}}

\hypertarget{section-3}{%
\section{4}\label{section-3}}

\bibverse{1} Then the messenger who talked with me came again and woke
me, as someone who is woken from sleep. \bibverse{2} He said to me,
``What do you see?'' I said, ``I see a candlestick, all of gold, with a
bowl on its top, and seven lamps on it. There are seven pipes to each of
the lamps, which are on its top, \bibverse{3} and two olive trees next
to it, one on the right side of the bowl, and the other on its left
side.'' \bibverse{4} I said to the messenger who talked with me, ``What
are these, my lord?'' \bibverse{5} Then the messenger who talked with me
answered, ``Don't you know what these are?'' And I said. ``No, my
lord.'' \bibverse{6} He said to me, ``This is the message of the Lord
regarding Zerubbabel: Not by might, nor by power, but by my spirit, says
the Lord of hosts, \bibverse{7} will I make the great mountain before
Zerubbabel a plain. He will bring forth the top stones with shoutings
of, `Grace, grace, to it.'\,'' \bibverse{8} Moreover this message of the
Lord came to me: \bibverse{9} The hands of Zerubbabel have laid the
foundation of this house. His hands will also finish it. Then you will
know that the Lord of hosts has sent me to you. \bibverse{10} For who
has despised the day of small things? They will rejoice, and will see
the plumb line in the hand of Zerubbabel. (The seven lamps are the eyes
of the Lord that search all around the world.) \bibverse{11} Again I
asked, ``What are these two olive trees on the right side of the
candlestick and on its left side?'' \bibverse{12} Again, ``What are
these two branches of the olive trees, which are pouring out the golden
oil through the two golden pipes?'' \bibverse{13} He answered me,
``Don't you now what these are?'' And I said, ``No, my lord.''
\bibverse{14} Then he explained, ``They are the two anointed ones who
attend the Lord of all the earth.''

\hypertarget{the-flying-scroll}{%
\subsection{The Flying Scroll}\label{the-flying-scroll}}

\hypertarget{section-4}{%
\section{5}\label{section-4}}

\bibverse{1} Then again I looked up and saw a flying scroll.
\bibverse{2} He said to me, ``What do you see?'' I answered, ``I see a
flying scroll; twenty cubits long, ten cubits wide.'' \bibverse{3} Then
he said to me, ``This is the curse that goes over the whole land. Every
thief will be banished according to the writing on one side and everyone
who lies under oath shall be banished according to the writing on the
other side. \bibverse{4} I have sent it out, says the Lord of hosts, and
it will enter into the house of the thief, and into the house of anyone
who swears falsely by my name. It will remain in the midst of his house
and will consume it with its timber and its stones.'' \#\# The Woman in
an Ephah

\bibverse{5} Then the messenger who talked with me came forward, and
said to me, ``Look up and see what is now coming.'' \bibverse{6} I said.
``What is it?'' He said, ``It is a measuring basket, full of the guilt
of the land.'' \bibverse{7} A round leaden cover was lifted up, and
there was a woman sitting in the middle of the basket. \bibverse{8} He
said, ``This is Wickedness.'' He thrust her down into the midst of the
basket and he pushed the round leaden cover back on its mouth.
\bibverse{9} Then I looked up and saw two women, and the wind was in
their wings. They had wings like the wings of a stork and they lifted up
the basket between them. \bibverse{10} Then I said to the messenger who
talked with me, ``Where are they taking the basket?'' \bibverse{11} He
said to me, ``To build her a house in the land of Shinar. When it is
prepared, they will rest the basket there.''

\hypertarget{the-war-chariots-of-the-lord}{%
\subsection{The War-Chariots of the
Lord}\label{the-war-chariots-of-the-lord}}

\hypertarget{section-5}{%
\section{6}\label{section-5}}

\bibverse{1} Again I looked up, and saw four chariots come from between
two mountains. The mountains were mountains of bronze. \bibverse{2} The
first chariot had red horses. The second chariot had black horses,
\bibverse{3} the third chariot white horses, and the fourth chariot
dappled horses. \bibverse{4} Then I said to the messenger who talked
with me, ``What are these, my lord?'' \bibverse{5} The messenger
answered me, ``These are the four winds of heaven going out after
standing before the Lord of all the earth. \bibverse{6} The black horses
are going towards the north country and the white towards the west and
the dappled towards the south country.'' \bibverse{7} The strong horses
came out, and were impatient to patrol the earth. So he said, ``Go,
patrol the earth.'' and they did so. \bibverse{8} Then he cried out to
me, ``The horses that went to the north country will quiet my spirit in
the north country.'' \#\# Preparations for the Crowning of Zerubbabel

\bibverse{9} Now this message of the Lord came to me: \bibverse{10} Take
the gifts from the exiles Heldai, Tobijah and Jedaiah who have come from
Babylon and go at once to Josiah the son of Zephaniah. \bibverse{11}
Take their silver and gold and make a crown and set it on the head of
the high priest, Joshua son of Jehozadak.

\bibverse{12} Tell him: ``Thus says the Lord of hosts: Behold, the man
whose name is the Branch. He will branch out from his place and rebuild
the temple of the Lord. \bibverse{13} He will rebuild the temple and
will sit and rule upon his throne, receiving the honour due to a king. A
priest will stand by his throne and they will work together in peace and
harmony. \bibverse{14} The crown will be a memorial in the temple of the
Lord for Heldai and Tobijah and Jedaiah, and Josiah the son of
Zephaniah. \bibverse{15} People from far away will come and build in the
temple of the Lord and you will know that the Lord of hosts has sent me
to you. This will happen if you diligently obey the voice of the Lord
your God.''

\hypertarget{the-lords-real-demands}{%
\subsection{The Lord's Real Demands}\label{the-lords-real-demands}}

\hypertarget{section-6}{%
\section{7}\label{section-6}}

\bibverse{1} In the fourth year of the reign of King Darius, on the
fourth day of Chislev, the ninth month, the message of the Lord came to
Zechariah. \bibverse{2} The city of Bethel had sent Sharezer and
Regem-melech and their men, to entreat the favour of the Lord,
\bibverse{3} and to ask the priests of the house of the Lord of hosts,
and the prophets ``Should I continue to mourn in the fifth month,
separating myself, as I have done for many years?'' \bibverse{4} Then
this message of the Lord of hosts came to me: \bibverse{5} Tell all the
people of the land and the priests, ``When you fasted and mourned in the
fifth and in the seventh month, for these seventy years, did you really
fast for me? \bibverse{6} When you eat and when you drink, do you not
eat for yourselves, and drink for yourselves? \bibverse{7} Wasn't it
this which the Lord proclaimed by the earlier prophets, when Jerusalem
was inhabited and prosperous, and her towns surrounding her and the
Negreb and the Shephelah were inhabited?''

\bibverse{8} The Lord gave this message to Zechariah: \bibverse{9} The
Lord of hosts says, execute true judgements. Show kindness and mercy to
each other. \bibverse{10} Do not oppress the widow or the orphan, the
resident alien or the poor. Let none of you devise evil against another
in your hearts.

\bibverse{11} But they refused to listen, and turned a stubborn
shoulder, and stopped their ears so they wouldn't hear. \bibverse{12}
They made their hearts as hard as stone so they wouldn't accept the
teaching and the words that the Lord of hosts had sent by his spirit
through the earlier prophets. Great was the anger of the Lord of hosts:
\bibverse{13} I called and they would not hear, so they call and I do
not hear, says the Lord of hosts. \bibverse{14} I scattered them by a
whirlwind out amongst nations unknown to them. The land was left
desolate behind them, with no one crossing or returning, for they made
the pleasant land a desolation.

\hypertarget{promises-that-the-scattered-exiles-will-return-and-the-lords-rule-be-widely-acknowledged}{%
\subsection{Promises that the Scattered Exiles Will Return and the
Lord's Rule be Widely
Acknowledged}\label{promises-that-the-scattered-exiles-will-return-and-the-lords-rule-be-widely-acknowledged}}

\hypertarget{section-7}{%
\section{8}\label{section-7}}

\bibverse{1} Now this message of the Lord of hosts came to me:
\bibverse{2} The Lord of hosts says: I passionately care for Zion, and I
am fiercely protective of her. \bibverse{3} The Lord says, I have
returned to Zion, and will dwell in the midst of Jerusalem and Jerusalem
will be called, `The City of Truth' and the mountain of the Lord of
hosts, `The Holy Mountain.' \bibverse{4} The Lord of hosts says: Old men
and old women will sit again in the squares of Jerusalem, each leaning
on a staff because of old age. \bibverse{5} And the streets of the city
will be full of boys, and of girls playing in its squares. \bibverse{6}
The Lord of hosts says: Just because it seems impossible to the remnant
of this people in these days, is it also impossible for me? says the
Lord of hosts. \bibverse{7} The Lord of hosts says: I will rescue my
people from the land of the east and the land of the west. \bibverse{8}
I will bring them and they will live in Jerusalem, and they will be my
people, and I will be their God, in truth and righteousness.

\bibverse{9} The Lord of hosts says: Let your hands be strong, you who
hear in these days the words of the prophets spoken when the foundations
for the house of the Lord of hosts were laid. \bibverse{10} For before
those days there were no wages for people or animals. No one could go
out or return safe from enemies, because I set each against their
neighbour. \bibverse{11} But now I will treat the remnant of the people
differently from former days, says the Lord of hosts. \bibverse{12}
Peace will be sown. The vine will bear its fruit and the ground will
yield its produce, and the skies will give their dew. I will give the
remnant of this people all these things. \bibverse{13} Just as you were
accursed amongst the nations, house of Judah and house of Israel, so I
will save you and you will be a blessing. Do not be afraid, but let your
hands be strong.

\bibverse{14} For the Lord of hosts says: As I planned to do evil to you
when your ancestors provoked me to anger, says the Lord of hosts, and I
did not relent, \bibverse{15} so again have I planned in these days to
do good to Jerusalem and to the house of Judah. Do not be afraid.

\bibverse{16} These are the things that you should do: Speak the truth
to each other. Render peaceful decisions in your gates. \bibverse{17}
Let none of you devise evil in your heart against one another. Do not
love perjury. All these are things I hate, says the Lord.

\bibverse{18} And this message of the Lord of hosts came to me:
\bibverse{19} The Lord of hosts says: The fast of the fourth month, and
the fast of the fifth, and the fast of the seventh, and the fast of the
tenth, will be festivals of joy and gladness, and cheerful feasts for
the house of Judah. Therefore love truth and peace.

\bibverse{20} The Lord of hosts says: Peoples, and the inhabitants of
many cities will come, \bibverse{21} and the inhabitants of one city
will go to another, saying, ``Let us go speedily to entreat the favour
of the Lord, and to seek the Lord of hosts; I will go also.''
\bibverse{22} Many peoples and strong nations will come to seek the Lord
of hosts in Jerusalem, and to entreat the favour of the Lord.
\bibverse{23} The Lord of hosts says: In those days ten people from all
languages and nations will take hold of the robe of a Jew, saying, ``We
will go with you, for we have heard that God is with you.''

\hypertarget{the-conquest-of-israels-foes}{%
\subsection{The Conquest of Israel's
Foes}\label{the-conquest-of-israels-foes}}

\hypertarget{section-8}{%
\section{9}\label{section-8}}

\bibverse{1} An oracle: The message of the Lord concerns the land of
Hadrach, it rests upon Damascus. For the capital of Aram belongs to the
Lord, as the tribes of Israel do, \bibverse{2} and also Hamath which
borders on it. Tyre and Sidon, for all their wisdom. \bibverse{3} Tyre
built for herself a fortress. Heaped up silver as dust, and gold like
the dirt of the streets. \bibverse{4} But now the Lord will dispossess
her, cast down her rampart into the sea. She will be consumed in fire.
\bibverse{5} Ashkelon will see and fear, and Gaza writhe in anguish,
Ekron, also for her hope will be put to shame. The king will perish from
Gaza, Ashkelon will remain uninhabited. \bibverse{6} A foreign people
will settle in Ashdod, and I will humble the pride of the Philistines.
\bibverse{7} I will take their blood from their mouth, and their
abominations from between their teeth. They also will be a remnant for
our God, like a clan in Judah, and Ekron will be like the Jebusites.
\bibverse{8} But I will encamp as a guard about my house, so that none
will pass through or return, and no oppressor will pass through them
again, for now I have seen with my own eyes. \bibverse{9} Rejoice
greatly, daughter of Zion! Shout aloud, daughter of Jerusalem! Look,
your king comes to you. Vindicated and victorious is he, humble, and
riding on a donkey, on a colt, the foal of a donkey. \bibverse{10} He
will cut off chariots from Ephraim, and war horses from Jerusalem. The
battle bow will be cut off, and he will speak peace to the nations. His
rule will be from sea to sea, from the river to the ends of the earth.
\bibverse{11} And you, too --- because of your blood covenant, I have
set free your prisoners from the waterless pit. \bibverse{12} To the
stronghold will the prisoners of hope return. This day I declare, I will
restore double to you. \bibverse{13} For I have bent Judah to me, like a
bow which I have filled with Ephraim like an arrow. I will urge your
sons, Zion, against the sons of Greece, I will make you like the sword
of a hero. \bibverse{14} Then the Lord will be seen above them and his
arrow will go forth like lightning. The Lord will blow a blast upon a
trumpet, and travel on the whirlwinds of the south. \bibverse{15} The
Lord of hosts will defend them. They will devour and tread down the
slingstones, they will drink their blood like wine, they will be full
like a bowl, like the crevices of the altar. \bibverse{16} And the Lord
their God will save them on that day, as the flock of his people, as the
jewel of a crown shining in his land. \bibverse{17} How good and how
beautiful will it be! Corn will make the young men flourish and new wine
the young women!

\hypertarget{victory-and-restoration-for-the-lords-betrayed-but-loyal-people}{%
\subsection{Victory and Restoration for the Lord's Betrayed but Loyal
People}\label{victory-and-restoration-for-the-lords-betrayed-but-loyal-people}}

\hypertarget{section-9}{%
\section{10}\label{section-9}}

\bibverse{1} Ask from the Lord rain, at the time of spring rain, from
the Lord who forms the storm clouds, and the showers of rain he gives to
you, and green in the field. \bibverse{2} But the household gods speak
nonsense, and the diviners see lies, and relate idle dreams. In vain
they offer comfort, so the people wander like sheep, They suffer because
there is no shepherd. \bibverse{3} My wrath is hot against the
shepherds, and on the leaders will I bring punishment. For the Lord
cares for his flock, the house of Judah, and will make them his splendid
war steeds. \bibverse{4} From Judah will come the cornerstone and the
stay, from them the battle bow, from them goes forth every commander.
\bibverse{5} Together they will trample on warriors, as on the dirt of
the streets in battle. They fight, for the Lord is with them, putting
the riders on horses to shame. \bibverse{6} I will make strong the house
of Judah, I will bring deliverance to the house of Joseph, I will bring
them back for I have compassion on them, they will be as if I had not
driven them away For I am the Lord their God and will answer them.
\bibverse{7} The Ephraimites will be like warriors, their hearts will
rejoice as with wine, their children will see and rejoice. Their hearts
will exult in the Lord. \bibverse{8} I will whistle for them, and gather
them, and they will be as many as they once were. \bibverse{9} I will
sow them amongst the peoples, but in distant lands they will remember
me, and they will bring up their children and come back. \bibverse{10} I
will bring them home from the land of Egypt, and from Assyria will I
gather them. Into the land of Gilead and Lebanon I will bring them,
until the land overflows with them. \bibverse{11} And when they pass
through the sea of distress, I will strike down the waves of the sea,
and all the deeps of the Nile will be dried up. The pride of Assyria
will be brought low and the sceptre of Egypt turned aside. \bibverse{12}
I will make them strong in the Lord, they will walk in his name, says
the Lord.

\hypertarget{fate-of-judahs-traitorous-rulers-and-the-ultimate-deliverance}{%
\subsection{Fate of Judah's Traitorous Rulers and the Ultimate
Deliverance}\label{fate-of-judahs-traitorous-rulers-and-the-ultimate-deliverance}}

\hypertarget{section-10}{%
\section{11}\label{section-10}}

\bibverse{1} Open your doors, Lebanon, so that fire may devour your
cedars. \bibverse{2} Wail, pine-tree, for the cedar is fallen. Wail,
oaks of Bashan, for the inaccessible forest has fallen. \bibverse{3}
Listen to the wailing of the shepherds! Their glory is destroyed. Listen
to the roaring of young lions! Blasted is the thick forest of the
Jordan.

\bibverse{4} The Lord said to me: shepherd the flock destined for
slaughter, \bibverse{5} whose possessors slaughter them, and hold
themselves not guilty! The people who sell them say, Blessed be the
Lord, for I am rich! And their own shepherds have no compassion for
them. \bibverse{6} I will no longer have pity on the people of this
earth, says the Lord. I will turn neighbour against neighbour, and put
everyone under the power of a king, and when the land is devastated I
will not rescue them. \bibverse{7} So I shepherded the flock of
slaughter for the sheep-merchants. I took two staffs: one I called Mercy
and the other I called Union. So I shepherded the sheep. \bibverse{8} I
set aside three of the shepherds in one month; for I was indignant
against them, and they also were displeased at me. \bibverse{9} I said,
I will not shepherd you. What is dead, let it die. What is set aside,
let it be set aside. And let those who are left devour one another's
flesh. \bibverse{10} And I took my staff, Mercy, and broke it, so as to
break my covenant which I had made with all the peoples.

\bibverse{11} In that day it was broken, and the sheep-merchants who
watched me knew that it was the Lord's message. \bibverse{12} And I said
to them, ``If it is good in your sight, give me my wage. If not,
don't.'' So they weighed out my wage, thirty pieces of silver.
\bibverse{13} And the Lord said to me, Cast it into the treasury --- the
precious wage that at which I was valued by them! So I took the thirty
pieces of silver and cast them into the house of the Lord, into the
treasury.

\bibverse{14} Then I broke my second staff called Union so as to
dissolve the brotherhood between Judah and Jerusalem. \bibverse{15} And
the Lord said to me: Take again the implements of a worthless shepherd.
\bibverse{16} For I am about to appoint a shepherd over the land. Those
who are thrust down he will not visit. Those who are scattered he will
not seek out. The wounded he will not heal, the sick he will not make
whole, but the flesh of the fat he will devour and even their hoofs he
will tear. \bibverse{17} Woe to my worthless shepherd, who deserts the
flock! May the sword strike his arm and his right eye. May his arm
completely wither. And his right eye be blinded.

\hypertarget{jerusalems-deliverance-from-heathen-attack}{%
\subsection{Jerusalem's Deliverance from Heathen
Attack}\label{jerusalems-deliverance-from-heathen-attack}}

\hypertarget{section-11}{%
\section{12}\label{section-11}}

\bibverse{1} An oracle. This is the message of the Lord about Israel.
The Lord who stretches out the heavens, and lays the foundation of the
earth, and forms the human spirit within people says: \bibverse{2} Look!
I am about to make Jerusalem a cup of drunkenness for all the
surrounding peoples. There will be a siege of Jerusalem. \bibverse{3} On
that day that I will make Jerusalem a stone to be lifted up by all the
peoples. All who lift it up will surely hurt themselves! And all the
nations of the earth will be gathered together against it. \bibverse{4}
On that day, says the Lord, I will strike every horse with panic and its
rider with madness. But over the house of Judah I will keep watch,
though I strike every horse belonging to the peoples with blindness.
\bibverse{5} And the chieftains of Judah will say to themselves, ``The
strength of the inhabitants is in the Lord of hosts their God.''
\bibverse{6} In that day I will make the chieftains of Judah like a pan
of fire in the woods, like a torch amongst sheaves, they will devour
right and left all the surrounding peoples. But Jerusalem will abide on
its own site. \bibverse{7} And the Lord will first give victory to the
tents of Judah, so that the glory of the house of David, and of the
inhabitants of Jerusalem be not exalted above Judah. \bibverse{8} In
that day the Lord will protect the inhabitants of Jerusalem, and the
feeblest amongst them will in that day be like David, and the house of
David like God, like the messenger of the Lord before them. \bibverse{9}
And in that day I will seek to destroy all the nations who have come up
against Jerusalem. \bibverse{10} And I will pour out on the house of
David and the inhabitants of Jerusalem the spirit of pity and
compassion. They will look on him whom they have pierced and they will
lament for him as one laments for an only son. They will bitterly grieve
for him as one grieves for the first-born. \bibverse{11} In that day
mourning will be as great in Jerusalem as the mourning for Hadad-rimmon
in the plain of Megiddo. \bibverse{12} And the land will mourn, each
family by itself: the family of the house of David by itself, and their
wives by themselves, and the family of the house of Nathan by itself,
and their wives by themselves, \bibverse{13} and the family of the house
of Levi by itself, and their wives by themselves, the family of the
Shimeites by itself, and their wives by themselves, \bibverse{14} and
all the families who are left, each by itself, and their wives by
themselves.

\hypertarget{section-12}{%
\section{13}\label{section-12}}

\bibverse{1} On that day a fountain will be opened for the house of
David, and for the inhabitants of Jerusalem, for sin and for
uncleanness. \bibverse{2} And it will be on that day, says the Lord of
hosts, I will cut off the names of the idols from the land and they will
be remembered no more. And the prophets and the unclean spirit I will
banish from the land. \bibverse{3} If any man prophesy any more, his
father and his mother who bore him, will say to him, ``You will not
live, for you speak falsehood in the name of the Lord.'' And his father
and his mother who bore him will stab him through when he is seized by
the prophetic frenzy. \bibverse{4} And it will be in that day that the
prophets will be ashamed, each of his vision, and will not wear the
hairy mantle in order to deceive. \bibverse{5} Each will say, ``I am not
a prophet, a tiller of the ground am I, for the ground is my possession
from my youth.'' \bibverse{6} And they will say to him, ``What are these
scars on your back?'' And he will say, ``The scars which I received in
the house of my friends.'' \bibverse{7} Awake, sword, against my
shepherd, against the man who stands near to me, says the Lord of hosts.
I will strike the shepherd so that the sheep may be scattered, I will
turn my hand against the little ones. \bibverse{8} In all the land, says
the Lord, two-thirds in it will be cut off and die, but a third will be
left in it. \bibverse{9} I will bring the third into the fire, and melt
it as one melts silver, and test it as one tests gold. Then they will
call on my name, and I myself will answer them. I will say, ``They are
my people.'' They will say, ``The Lord is our God.''

\hypertarget{the-judgement-upon-the-heathen-and-the-exaltation-of-the-lord}{%
\subsection{The Judgement upon the Heathen and the Exaltation of the
Lord}\label{the-judgement-upon-the-heathen-and-the-exaltation-of-the-lord}}

\hypertarget{section-13}{%
\section{14}\label{section-13}}

\bibverse{1} A day is coming for the Lord, when your plundered
possessions will be divided while you watch. \bibverse{2} And all the
nations will gather to Jerusalem, to fight against it. The city will be
taken and the houses plundered and the women raped. Half of the city
will go into captivity, and the rest of the people left in the ruins.
\bibverse{3} Then the Lord will go forth and fight against these
nations, as once he fought in the day of battle. \bibverse{4} On that
day his feet will stand on the Mount of Olives, (which is opposite
Jerusalem, on the east). The Mount of Olives will be split into halves,
from east to west, by an exceedingly great valley; and half of the
mountain will slide northwards and half southwards. \bibverse{5} You
will escape through my valley --- the valley between the hills will
extend as far as Azel, and you will flee as you fled from before the
earthquake, in the days of Uzziah king of Judah. But the Lord your God
will come, and all the holy ones with him.

\bibverse{6} And in that day, there will be no heat nor cold nor frost,
\bibverse{7} but it will be constant day --- it is known to the Lord ---
with neither day nor night. Even at evening time there will be light.
\bibverse{8} And on that day living waters will flow from Jerusalem,
half of them to the eastern sea and half of them to the western sea; in
both summer and winter. \bibverse{9} The Lord will be king over all the
earth. On that day the Lord will be one, and his name one. \bibverse{10}
The land will be changed to plain, from Geba to Rimmon, south of
Jerusalem, but Jerusalem will be high and inhabited as it stands, from
the Benjamin Gate up to the place of the first gate, and from the Tower
of Hananel to the Corner Gate, and as far as the king's winepresses.
\bibverse{11} Jerusalem will be inhabited, for never again will it be
doomed to destruction, and its people will abide in security.
\bibverse{12} This is the plague with which the Lord will strike all the
peoples who array themselves against Jerusalem: he will make their flesh
rot while they stand on their feet, and their eyes will rot in their
sockets, and their tongues will rot in their mouth. \bibverse{13} On
that day a great panic sent by the Lord will fall on them. Everyone will
seize their neighbours and attack them. \bibverse{14} Even Judah will
fight at Jerusalem, and the wealth of the surrounding nations will be
gathered up --- gold and silver and piles of clothing. \bibverse{15}
Plague will fall upon the horses, mules, camels, and asses, and on all
the animals in all those camps.

\bibverse{16} All who are left of all the nations which fought against
Jerusalem will come up from year to year to worship the King, the Lord
of hosts, and to keep the pilgrim-feast of tabernacles. \bibverse{17}
Whoever of all the peoples of the earth will not come up to Jerusalem to
worship the King, the Lord of hosts, on them there will be no rain.
\bibverse{18} If the family of Egypt does not go up nor enter in, on
them also will come the plague with which the Lord will strike the
nations. \bibverse{19} This will be the punishment for the sin of Egypt
and the punishment for the sin of all nations which do not come up to
keep the feast of the tabernacles.

\bibverse{20} On that day there will be inscribed upon the bells of the
horses, `Holy to the Lord' and the pots in the house of the Lord will be
as holy as the sacrificial bowls before the altars. \bibverse{21} Every
pot in Jerusalem and in Judah will be holy to the Lord of hosts and all
who sacrifice will come and take of them and cook the sacrifices in
them. There will be no more traffickers in the house of the Lord of
hosts in that day.
