\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

\hypertarget{section}{%
\section{1}\label{section}}

\bibverse{1} To the church of God in Corinth, to those who have been
consecrated by union with Christ Jesus and called to become his people,
and also to all, wherever they may be, who invoke the name of our Lord
Jesus Christ --- their Master and ours, \bibverse{2} from Paul, who has
been called to be an apostle of Jesus Christ by the will of God, and
from Sosthenes, our fellow follower of the Lord.

\bibverse{3} May God, our Father, and the Lord Jesus Christ bless you
and give you peace. \bibverse{4} I always thank God about you for the
blessing bestowed on you in Christ Jesus. \bibverse{5} For through union
with him you were enriched in every way --- in your power to preach, and
in your knowledge of the truth; \bibverse{6} and so became yourselves a
confirmation of my testimony to the Christ. \bibverse{7} There is no
gift in which you are deficient, while waiting for the appearing of our
Lord Jesus Christ. \bibverse{8} And God himself will strengthen you to
the end, so that at the day of our Lord Jesus Christ you may be found
blameless. \bibverse{9} God will not fail you, and it is he who called
you into communion with his Son, Jesus Christ, our Lord. \#\# The State
of the Church at Corinth

\bibverse{10} But I appeal to you, my friends, by the name of our Lord
Jesus Christ, to agree in what you profess, and not to allow divisions
to exist amongst you, but to be united --- of one mind and of one
opinion. \bibverse{11} For I have been informed, my friends, by the
members of Chloe's household, that party feeling exists amongst you.
\bibverse{12} I mean this: that every one of you says either `I follow
Paul,' or `I Apollos,' or `I Cephas,' or `I Christ.' \bibverse{13} You
have torn the Christ in pieces! Was it Paul who was crucified for you?
Or were you baptised into the faith of Paul? \bibverse{14} I am thankful
that I did not baptise any of you except Crispus and Gaius,
\bibverse{15} so that no one can say that you were baptised into my
faith. \bibverse{16} I baptised also the household of Stephanas. I do
not know that I baptised anyone else. \bibverse{17} My mission from
Christ was not to baptise, but to tell the good news; not, however, in
the language of philosophy, in case the cross of the Christ should be
robbed of its meaning. \bibverse{18} The message of the cross is indeed
mere folly to those who are in the path to ruin, but to us who are in
the path of salvation it is the power of God. \bibverse{19} For
scripture says --- `I will bring the philosophy of the philosophers to
naught, and the shrewdness of the shrewd I will bring to nothing.'
\bibverse{20} Where is the philosopher? Where the teacher of the Law?
Where the disputant of today? Has not God shown the world's philosophy
to be folly? \bibverse{21} For since the world, in God's wisdom, did not
by its philosophy learn to know God, God saw fit, by the `folly' of our
proclamation, to save those who believe in Christ! \bibverse{22} While
Jews ask for miraculous signs, and Greeks study philosophy,
\bibverse{23} we are proclaiming Christ crucified! --- to the Jews an
obstacle, to the Gentiles mere folly, \bibverse{24} but to those who
have received the call, whether Jews or Greeks, Christ, the power of God
and the wisdom of God! \bibverse{25} For God's `folly' is wiser than
people, and God's `weakness' is stronger than people.

\bibverse{26} Look at the facts of your call, friends. There are not
many amongst you who are wise, as people reckon wisdom, not many who are
influential, not many who are high-born; \bibverse{27} but God chose
what the world counts foolish to put its wise to shame, and God chose
what the world counts weak to put its strong to shame, \bibverse{28} and
God chose what the world counts poor and insignificant --- things that
to it are unreal --- to bring its `realities' to nothing, \bibverse{29}
so that in his presence no one should boast. \bibverse{30} But you, by
your union with Christ Jesus, belong to God; and Christ, by God's will,
became not only our wisdom, but also our righteousness, holiness, and
deliverance, \bibverse{31} so that --- in the words of scripture ---
`Whoever boasts should boast of the Lord!'

\hypertarget{section-1}{%
\section{2}\label{section-1}}

\bibverse{1} For my own part, friends, when I came to you, it was with
no display of eloquence or philosophy that I came to tell the hidden
purpose of God; \bibverse{2} for I had determined that, while with you,
I would forget everything Jesus Christ --- and him crucified!
\bibverse{3} Indeed, when I came amongst you, I was weak, and full of
fears, and in great anxiety. \bibverse{4} My message and my proclamation
were not delivered in the persuasive language of philosophy, but were
accompanied by the manifestation of spiritual power, \bibverse{5} so
that your faith should be based, not on the human wisdom, but on the
power of God.

\bibverse{6} Yet there is a philosophy that we teach to those whose
faith is matured, but it is not the philosophy of today, or of the
leaders of today --- whose downfall is at hand. \bibverse{7} No, it is a
divine philosophy that we teach, one concerned with the hidden purpose
of God --- that long hidden philosophy which God, before time began,
destined for our glory. \bibverse{8} This philosophy is not known to any
of the leaders of today; for, had they known it, they would not have
crucified our glorified Lord. \bibverse{9} It is what scripture speaks
of as --- `What eye never saw, nor ear ever heard, what never entered
people's minds --- even all that God has prepared for those who love
him.' \bibverse{10} Yet to us God revealed it through his Spirit; for
the Spirit fathoms all things, even the inmost depths of God's being.
\bibverse{11} For what person is there who knows what a person is,
except the person's own spirit within them? So, also, no one comprehends
what God is, except the Spirit of God. \bibverse{12} And as for us, it
is not the spirit of the world that we have received, but the Spirit
that comes from God, so that we may realise the blessings given to us by
him. \bibverse{13} And we speak of these gifts, not in language taught
by human philosophy, but in language taught by the Spirit, explaining
spiritual things in spiritual words. \bibverse{14} The merely
intellectual person rejects the teaching of the Spirit of God; for to
them it is mere folly; they cannot grasp it, because it is to be
understood only by spiritual insight. \bibverse{15} But the person with
spiritual insight is able to understand everything, although they
themselves are understood by no one. \bibverse{16} For who has so
comprehended the mind of the Lord as to be able to instruct him? We,
however, have the mind of Christ.

\hypertarget{section-2}{%
\section{3}\label{section-2}}

\bibverse{1} But I, my friends, could not speak to you as people with
spiritual insight, but only as worldly-minded --- mere infants in the
faith of Christ. \bibverse{2} I fed you with milk, not with solid food,
for you were not then able to take it.

No, and even now you are not able; you are still worldly. \bibverse{3}
While there exist amongst you jealousy and party feeling, is it not true
that you are worldly, and are acting merely as other people do?
\bibverse{4} When one says `I follow Paul,' and another `I follow
Apollos,' aren't you like other people? \bibverse{5} What, I ask, is
Apollos? Or what is Paul? Servants through whom you were led to accept
the faith; and that only as the Lord helped each of you. \bibverse{6} I
planted, and Apollos watered, but it was God who caused the growth.
\bibverse{7} Therefore neither the one who plants, nor the one who
waters, counts for anything, but only God who causes the growth.
\bibverse{8} In this the person who plants and the person who waters are
one; yet each will receive their own reward in proportion to their own
labour. \bibverse{9} For we are God's fellow workers; you are God's
harvest field, God's building.

\bibverse{10} In fulfilment of the charge which God had entrusted to me,
I laid the foundation like a skilful master; but someone else is now
building on it. Let everyone take care how they build; \bibverse{11} for
no one can lay any other foundation than the one already laid --- Jesus
Christ. \bibverse{12} Whatever is used by those who build on this
foundation, whether gold, silver, costly stones, wood, hay, or straw,
\bibverse{13} the quality of each man's work will become known, for the
day will make it plain; because that day is to be ushered in with fire,
and the fire itself will test the quality of every man's work.
\bibverse{14} If anyone's work, which they have built on that
foundation, still remains, they will gain a reward. \bibverse{15} If
anyone's work is burnt up, they will suffer loss; though they themselves
will escape, but only as one who has passed through fire.

\bibverse{16} Don't you know that you are God's Temple, and that God's
Spirit has his home in you? \bibverse{17} If any one destroys the Temple
of God, God will destroy them; for the Temple of God is sacred, and so
also are you. \bibverse{18} Let no one deceive himself. If any one
amongst you imagines that, as regards this world, they are wise, they
should become a `fool,' that they may become wise. \bibverse{19} For in
God's sight this world's wisdom is folly. Scripture tells of --- `One
who catches the wise in their own craftiness,' \bibverse{20} And it says
again --- `The Lord sees how fruitless are the deliberations of the
wise.' \bibverse{21} Therefore let no one boast about people; for all
things are yours --- \bibverse{22} Whether Paul, or Apollos, or Cephas,
or the world, or life, or death, or the present, or the future --- all
things are yours! \bibverse{23} But you are Christ's and Christ is
God's.

\hypertarget{section-3}{%
\section{4}\label{section-3}}

\bibverse{1} Let people look on us as Christ's servants, and as stewards
of the hidden truths of God. \bibverse{2} Now what we look for in
stewards is that they should be trustworthy. \bibverse{3} But it weighs
very little with me that I am judged by you or by any human tribunal.
No, I do not even judge myself; \bibverse{4} for, though I am conscious
of nothing against myself, that does not prove me innocent. It is the
Lord who is my judge. \bibverse{5} Therefore do not pass judgement
before the time, but wait until the Lord comes. He will throw light on
what is now dark and obscure, and will reveal the motives in people's
minds; and then everyone will receive due praise from God.

\bibverse{6} All this, friends, I have, for your sakes, applied to
Apollos and myself, so that, from our example, you may learn to observe
the precept --- `Keep to what is written,' that none of you may speak
boastfully of one teacher to the disparagement of another. \bibverse{7}
For who makes any one of you superior to others? And what have you that
was not given you? But if you received it as a gift, why do you boast as
if you had not? \bibverse{8} Are you all so soon satisfied? Are you so
soon rich? Have you begun to reign without us? Would indeed that you
had, so that we also might reign with you! \bibverse{9} For, as it seems
to me, God has exhibited us, the apostles, last of all, as people doomed
to death. We are made a spectacle to the universe, both to angels and to
people! \bibverse{10} We, for Christ's sake, are fools, but you, by your
union with Christ, are people of discernment. We are weak, but you are
strong. You are honoured, but we are despised. \bibverse{11} To this
very hour we go hungry, thirsty, and naked; we are beaten; we are
homeless; \bibverse{12} we work hard, toiling with our own hands. We
meet abuse with blessings, we meet persecution with endurance,
\bibverse{13} we meet slander with gentle appeals. We have been treated
as the scum of the earth, the vilest of the vile, to this very hour.

\bibverse{14} It is with no wish to shame you that I am writing like
this; but to warn you as my own dear children. \bibverse{15} Though you
may have thousands of instructors in the faith of Christ, yet you have
not many fathers. It was I who, through union with Christ Jesus, became
your father by means of the good news. \bibverse{16} Therefore I entreat
you --- Follow my example. \bibverse{17} This is my reason for sending
Timothy to you. He is my own dear faithful child in the Master's
service, and he will remind you of my methods of teaching the faith of
Christ Jesus --- methods which I follow everywhere in every church.

\bibverse{18} Some, I hear, are puffed up with pride, thinking that I am
not coming to you. \bibverse{19} But come to you I will, and that soon,
if it please the Lord; and then I will find out, not what words these
people use who are so puffed up, but what power they possess;
\bibverse{20} for the kingdom of God is based, not on words, but on
power. \bibverse{21} What do you wish? Am I to come to you with a rod,
or in a loving and gentle spirit?

\hypertarget{section-4}{%
\section{5}\label{section-4}}

\bibverse{1} There is a wide-spread report respecting a case of
immorality amongst you, and that, too, of a kind that does not occur
even amongst the Gentiles --- a man, I hear, is living with his father's
wife! \bibverse{2} Instead of grieving over it and taking steps for the
expulsion of the man who has done this thing, is it possible that you
are still puffed up? \bibverse{3} For I myself, though absent in body,
have been present with you in spirit, and in the name of our Lord Jesus
I have already passed judgement, just as if I had been present, on the
man who has acted in this way. \bibverse{4} I have decided --- having
been present in spirit at your meetings, when the power of the Lord
Jesus was with us --- \bibverse{5} to deliver such a man as this over to
Satan, that what is sensual in him may be destroyed, so that his spirit
may be saved at the day of the Lord. \bibverse{6} Your boasting is
unseemly. Don't you know that even a little leaven leavens all the
dough? \bibverse{7} Get rid entirely of the old leaven, so that you may
be like new dough --- free from leaven, as in truth you are. For our
passover lamb is already sacrificed --- Christ himself; \bibverse{8}
therefore let us keep our festival, not with the leaven of former days,
nor with the leaven of vice and wickedness, but with the unleavened
bread of sincerity and truth.

\bibverse{9} I told you, in my letter, not to associate with immoral
people --- \bibverse{10} not, of course, meaning people of the world who
are in immoral, or who are covetous and grasping, or who worship idols;
for then you would have to leave the world altogether. \bibverse{11}
But, as things are, I say that you are not to associate with anyone who,
although a follower of Christ in name, is immoral, or covetous, or an
idolater, or abusive, or a drunkard, or grasping --- no, not even to sit
down to eat with such people. \bibverse{12} What have I to do with
judging those outside the church? Is it not for you to judge those who
are within the church, \bibverse{13} while God judges those who are
outside? Put away the wicked from amongst you.

\hypertarget{section-5}{%
\section{6}\label{section-5}}

\bibverse{1} Can it be that, when one of you has a dispute with another,
they dare to have their case tried before the heathen, instead of before
Christ's people? \bibverse{2} Don't you know that Christ's people will
try the world? And if the world is to be tried by you, are you unfit to
try the most trivial cases? \bibverse{3} Don't you know that we are to
try angels --- to say nothing of the affairs of this life? \bibverse{4}
Why, then, if you have cases relating to the affairs of this life, do
you set to try them those who carry no weight with the church? To your
shame I ask it. \bibverse{5} Can it be that there is not one amongst you
wise enough to decide between two of their fellow followers?
\bibverse{6} Must a follower sue a fellow follower? In front of
unbelievers? \bibverse{7} To begin with, it is undoubtedly a loss to you
to have lawsuits with one another. Why not rather let yourselves be
wronged? Why not rather let yourselves be cheated? \bibverse{8} Instead
of this, you wrong and cheat others yourselves --- yes, even other
followers! \bibverse{9} Don't you know that wrong-doers will have no
share in God's kingdom? Do not be deceived. No one who is immoral, or an
idolater, or an adulterer, or licentious, or a sexual pervert,
\bibverse{10} or a thief, or covetous, or a drunkard, or abusive, or
grasping, will have any share in God's kingdom. \bibverse{11} Such some
of you used to be; but you washed yourselves clean. You became Christ's
people! You were pronounced righteous through the name of our Lord Jesus
Christ, and through the Spirit of our God! \bibverse{12} Everything is
allowable for me! Yes, but everything is not profitable. Everything is
allowable for me! Yes, but for my part, I will not let myself be
enslaved by anything. \bibverse{13} Food exists for the stomach, and the
stomach for food; but God will put an end to both the one and the other.
The body, however, exists, not for immorality, but for the Lord, and the
Lord for the body; \bibverse{14} and, as God has raised the Lord, so he
will raise up us also by the exercise of his power. \bibverse{15} Don't
you know that your bodies are Christ's members? Am I, then, to take the
members that belong to the Christ and make them the members of a
prostitute? Heaven forbid! \bibverse{16} Or don't you know that a man
who unites himself with a prostitute is one with her in body (for `the
two,' it is said, `will become one'); \bibverse{17} while a man who is
united with the Lord is one with him in spirit? \bibverse{18} Shun all
immorality. Every other sin that people commit is something outside the
body; but an immoral person sins against their own body. \bibverse{19}
Again, don't you know that your body is a shrine of the Holy Spirit that
is within you -- the Spirit which you have from God? \bibverse{20}
Moreover, you are not your own masters; you were bought, and the price
was paid. Therefore, honour God in your bodies.

\hypertarget{answers-to-questions-asked-by-the-church-at-corinth}{%
\subsection{Answers to Questions Asked by the Church at
Corinth}\label{answers-to-questions-asked-by-the-church-at-corinth}}

\hypertarget{section-6}{%
\section{7}\label{section-6}}

\bibverse{1} With reference to the subjects about which you wrote to me:
It is good for a man to remain single. \bibverse{2} But, owing to the
prevalence of immorality, I advise every man to have his own wife, and
every woman her husband. \bibverse{3} A husband should give his wife her
due, and a wife her husband. \bibverse{4} It is not the wife, but the
husband, who exercises power over her body; and so, too, it is not the
husband, but the wife, who exercises power over his body. \bibverse{5}
Do not deprive each other of what is due --- unless it is only for a
time and by mutual consent, so that your minds may be free for prayer
until you again live as man and wife --- otherwise Satan might take
advantage of your want of self-control and tempt you. \bibverse{6} I say
this, however, as a concession, not as a command. \bibverse{7} I should
wish everyone to be just what I am myself. But everyone has his own gift
from God --- one in one way, and one in another. \bibverse{8} My advice,
then, to those who are not married, and to widows, is this: It would be
good for them to remain as I am myself. \bibverse{9} But, if they cannot
control themselves, they should marry, for it is better to marry than to
be consumed with passion. \bibverse{10} To those who are married my
direction is --- yet it is not mine, but the Master's --- that a woman
is not to leave her husband \bibverse{11} (If she has done so, she
should remain as she is, or else be reconciled to her husband) and also
that a man is not to divorce his wife. \bibverse{12} To all others I say
--- I, not the Master --- If a follower of the Lord is married to a
woman, who is an unbeliever but willing to live with him, he should not
divorce her; \bibverse{13} and a woman who is married to a man, who is
an unbeliever but willing to live with her, should not divorce her
husband. \bibverse{14} For, through his wife, the husband who is an
unbeliever has become associated with Christ's people; and the wife who
is an unbeliever has become associated with Christ's people through the
Lord's follower whom she has married. Otherwise your children would be
`defiled,' but, as it is, they belong to Christ's people. \bibverse{15}
However, if the unbeliever wishes to be separated, let them be so. Under
such circumstances neither is bound; God has called you to live in
peace. \bibverse{16} How can you tell, wife, whether you may not save
your husband? And how can you tell, husband, whether you may not save
your wife?

\bibverse{17} In any case, a person should continue to live in the
condition which the Lord has allotted to them, and in which they were
when God called them. This is the rule that I lay down in every church.
\bibverse{18} Was a man already circumcised when he was called? Then he
should not efface his circumcision. Has a man been called when
uncircumcised? Then he should not be circumcised. \bibverse{19}
Circumcision is nothing; the want of it is nothing; but to keep the
commands of God is everything. \bibverse{20} Let everyone remain in that
condition of life in which they were when the call came to them.
\bibverse{21} Were you a slave when you were called? Do not let that
trouble you. No, even if you are able to gain your freedom, still do
your best. \bibverse{22} For the person who was a slave when they were
called to the master's service is the Master's freedman; so, too, the
person who was free when called is Christ's slave. \bibverse{23} You
were bought, and the price was paid. Do not let yourselves become slaves
to people. \bibverse{24} Friends, let everyone remain in the condition
in which they were when they were called, in close communion with God.
\bibverse{25} With regard to unmarried women, I have no command from the
Master to give you, but I tell you my opinion, and the Master in his
mercy has made me worthy to be trusted. \bibverse{26} I think, then,
that, in view of the time of suffering that has now come upon us, what I
have already said is best --- that a man should remain as he is.
\bibverse{27} Are you married to a wife? Then do not seek to be
separated. Are you separated from a wife? Then do not seek for a wife.
\bibverse{28} Still, if you should marry, that is not wrong; nor, if a
young woman marries, is that wrong. But those who marry will have much
trouble to bear, and my wish is to spare you. \bibverse{29} What I mean,
friends, is this --- the time is short. Meanwhile, let those who have
wives live as if they had none, \bibverse{30} those who are weeping as
if not weeping, those who are rejoicing as if not rejoicing, those who
buy as if not possessing, \bibverse{31} and those who use the good
things of the world as using them sparingly; for this world as we see it
is passing away. \bibverse{32} I want you to be free from anxiety. The
unmarried man is anxious about the Master's cause, desiring to please
him; \bibverse{33} while the married man is anxious about worldly
matters, desiring to please his wife; \bibverse{34} and so his interests
are divided. Again, the unmarried woman, whether she is old or young, is
anxious about the Master's cause, striving to be pure both in body and
in spirit, while the married woman is anxious about worldly matters,
desiring to please her husband. \bibverse{35} I say this for your own
benefit, not with any intention of putting a halter around your necks,
but in order to secure for the Master seemly and constant devotion, free
from all distraction.

\bibverse{36} If, however, a father thinks that he is not acting fairly
by his unmarried daughter, when she is past her youth, and if under
these circumstances her marriage ought to take place, he should act as
he thinks right. He is doing nothing wrong --- let the marriage take
place. \bibverse{37} On the other hand, a father, who has definitely
made up his mind, and is under no compulsion, but is free to carry out
his own wishes, and who has come to the decision, in his own mind, to
keep his unmarried daughter at home will be doing right. \bibverse{38}
In short, the one who consents to his daughter's marriage is doing
right, and yet the other will be doing better.

\bibverse{39} A wife is bound to her husband as long as he lives; but,
if the husband should pass to his rest, the widow is free to marry
anyone she wishes, provided he is a believer. \bibverse{40} Yet she will
be happier if she remains as she is --- in my opinion, for I think that
I also have the Spirit of God.

\hypertarget{section-7}{%
\section{8}\label{section-7}}

\bibverse{1} With reference to food that has been offered in sacrifice
to idols --- We are aware that all of us have knowledge! Knowledge
breeds conceit, while love builds up character. \bibverse{2} If someone
thinks that they know anything, they have not yet reached that knowledge
which they ought to have reached. \bibverse{3} On the other hand, if a
person loves God, they are known by God. \bibverse{4} With reference,
then, to eating food that has been offered to idols --- we are aware
that an idol is nothing in the world, and that there is no God but one.
\bibverse{5} Even supposing that there are so-called `gods' either in
heaven or on earth --- and there are many such `gods' and `lords' ---
\bibverse{6} Yet for us there is only one God, the Father, from whom all
things come (and for him we live), and one Lord, Jesus Christ, through
whom all things come (and through him we live). \bibverse{7} Still, it
is not everyone that has this knowledge. Some people, because of their
association with idols, continued down to the present time, eat the food
as food offered to an idol; and their consciences, while still weak, are
dulled. \bibverse{8} What we eat, however, will not bring us nearer to
God. We lose nothing by not eating this food, and we gain nothing by
eating it. \bibverse{9} But take care that this right of yours does not
become in any way a stumbling-block to the weak. \bibverse{10} For if
someone should see you who possess this knowledge, feasting in an idol's
temple, will not their conscience, if they are weak, become so hardened
that they, too, will eat food offered to idols? \bibverse{11} And so,
through this knowledge of yours, the weak person is ruined --- someone
for whose sake Christ died! \bibverse{12} In this way, by sinning
against your fellow followers of the Lord and injuring their
consciences, while still weak, you sin against Christ. \bibverse{13}
Therefore, if what I eat makes a follower of the Lord fall, rather than
make them fall, I will never eat meat again.

\hypertarget{section-8}{%
\section{9}\label{section-8}}

\bibverse{1} Am I not free? Am I not an apostle? Have I not seen our
Lord Jesus? Aren't you yourselves my work achieved in union with the
Lord? \bibverse{2} If I am not an apostle to others, yet at least I am
to you; for you are the seal that stamps me as an apostle in union with
the Lord. \bibverse{3} The defence that I make to my critics is this:
\bibverse{4} Haven't we a right to food and drink? \bibverse{5} Haven't
we a right to take a wife with us, if she is a Christian, as the other
apostles and the Master's brothers and Cephas all do? \bibverse{6} Or is
it only Barnabas and I who have no right to give up working for our
bread? \bibverse{7} Does anyone ever serve as a soldier at his own
expense? Does anyone plant a vineyard and not eat its produce? Or does
anyone look after a herd and not drink the milk? \bibverse{8} Am I, in
all this, speaking only from the human standpoint? Does not the Law also
say the same? \bibverse{9} For in the Law of Moses it is said --- `You
should not muzzle a bullock while it is treading out the grain.' Is it
the bullocks that God is thinking of? \bibverse{10} Or is not is said
entirely for our sakes? Surely it was written for our sakes, for the
ploughman ought not to plough, nor the thrasher to thrash, without
expecting a share of the grain. \bibverse{11} Since we, then, sowed
spiritual seed for you, is it too much that we should reap from you an
earthly harvest? \bibverse{12} If others share in this right over you,
don't we even more? Still we did not avail ourselves of this right. No,
we endure anything rather than impede the progress of the good news of
the Christ. \bibverse{13} Don't you know that those who do the work of
the Temple live on what comes from the Temple, and that those who serve
at the altar share the offerings with the altar? \bibverse{14} So, too,
the Master has appointed that those who tell the good news should get
their living from the good news. \bibverse{15} I, however, have not
availed myself of any of these rights. I am not saying this to secure
such an arrangement for myself; indeed, I would far rather die ---
Nobody will make my boast a vain one! \bibverse{16} If I tell the good
news, I have nothing to boast of, for I am compelled to do so. Woe is me
if I do not tell it! \bibverse{17} If I do this work willingly, I have a
reward; but, if unwillingly, I have been charged to perform a duty.
\bibverse{18} What is my reward, then? To present the good news free of
all cost, and so make but a sparing use of the rights which it gives me.

\bibverse{19} Although I was entirely free, yet, to win as many converts
as possible, I made myself everyone's slave. \bibverse{20} To the Jews I
became like a Jew, to win Jews. To those who are subject to Law I became
like a man subject to Law --- though I was not myself subject to Law ---
to win those who are subject to Law. \bibverse{21} To those who have no
Law I became like a man who has no Law --- not that I am free from God's
Law; no, for I am under Christ's Law --- to win those who have no law.
\bibverse{22} To the weak I became weak, to win the weak. I have become
all things to all people, so as at all costs to save some. \bibverse{23}
And I do everything for the sake of the good news, so that with them I
may share in its blessings.

\bibverse{24} Don't you know that on a racecourse, though all run, yet
only one wins the prize? Run in such a way that you may win.
\bibverse{25} Every athlete exercises self-restraint in everything;
they, indeed, for a crown that fades, we for one that is unfading.
\bibverse{26} I, therefore, run with no uncertain aim. I box --- not
like a man hitting the air. \bibverse{27} No, I bruise my body and make
it my slave, so that I, who have called others to the contest, will not
myself be rejected.

\hypertarget{section-9}{%
\section{10}\label{section-9}}

\bibverse{1} I want you to bear in mind, friends, that all our ancestors
were beneath the cloud, and all passed through the sea; \bibverse{2}
that in the cloud and in the sea they all underwent baptism as followers
of Moses; \bibverse{3} and that they all ate the same supernatural food,
\bibverse{4} and all drank the same supernatural water, for they used to
drink from a supernatural rock which followed them, and that rock was
the Christ. \bibverse{5} Yet with most of them God was displeased; for
they were struck down in the desert. \bibverse{6} Now these things
happened as warnings to us, to teach us not to long for evil things as
our ancestors longed. \bibverse{7} Do not become idolaters, as some of
them became. Scripture says --- `The people sat down to eat and drink,
and stood up to dance.' \bibverse{8} Nor let us act immorally, as some
of them acted, with the result that twenty-three thousand of them fell
dead in a single day. \bibverse{9} Nor let us try the patience of the
Lord too far, as some of them tried it, with the result that they were,
one after another, destroyed by the snakes. \bibverse{10} And do not
murmur, as some of them murmured, and so were destroyed by the angel of
death. \bibverse{11} These things happened to them by way of warning,
and were recorded to serve as a caution to us, in whose days the close
of the ages has come.

\bibverse{12} Therefore let the person who thinks that they stand take
care that they do not fall. \bibverse{13} No temptation has come upon
you that is not common to all humanity. God will not fail you, and he
will not allow you to be tempted beyond your strength; but, when he
sends the temptation, he will also provide the way of escape, so that
you may have strength to endure. \bibverse{14} Therefore, my dear
friends, shun the worship of idols. \bibverse{15} I speak to you as man
of discernment; form your own judgement about what I am saying.
\bibverse{16} In the cup of blessing which we bless, is not there a
sharing in the blood of the Christ? And in the bread which we break, is
not there a sharing in the body of the Christ? \bibverse{17} The bread
is one, and we, though many, are one body; for we all partake of that
one bread. \bibverse{18} Look at the people of Israel. Do not those who
eat the sacrifices share with the altar? \bibverse{19} What do I mean?
That an offering made to an idol, or the idol itself, is anything?
\bibverse{20} No; what I say is that the sacrifices offered by the
Gentiles are offered to demons and to a being who is no God, and I do
not want you to share with demons. \bibverse{21} You cannot drink both
the Cup of the Lord and the cup of demons. You cannot partake at the
Table of the Lord and at the table of demons. \bibverse{22} Or are we to
rouse the jealousy of the Lord? Are we stronger than he?

\bibverse{23} Everything is allowable! Yes, but everything is not
profitable. Everything is allowable! Yes, but everything does not build
up character. \bibverse{24} A person must not study their own interests,
but the interests of others.

\bibverse{25} Eat anything that is sold in the market, without making
inquiries to satisfy your scruples; \bibverse{26} for the earth, with
all that is in it, belongs to the Lord. \bibverse{27} If an unbeliever
invites you to his house and you consent to go, eat anything that is put
before you, without making inquiries to satisfy your scruples.
\bibverse{28} But, if anyone should say to you `This has been offered in
sacrifice to an idol,' then, for the sake of the speaker and his
scruples, do not eat it. \bibverse{29} I do not say your scruples, but
his. For why should the freedom that I claim be condemned by the
scruples of another? \bibverse{30} If, for my part, I take the food
thankfully, why should I be abused for eating that for which I give
thanks?

\bibverse{31} Whether, then, you eat or drink or whatever you do, do
everything to the honour of God. \bibverse{32} Do not cause offence
either to Jews or Greeks or to the church of God; \bibverse{33} for I,
also, try to please everybody in everything, not seeking my own
advantage, but do what is best for others, so that they may be saved.

\hypertarget{section-10}{%
\section{11}\label{section-10}}

\bibverse{1} Imitate me, as I myself imitate Christ. \bibverse{2} I
praise you, indeed, because you never forget me, and are keeping my
injunctions in mind, exactly as I laid them on you. \bibverse{3} But I
am anxious that you should understand that the Christ is the head of
every man, that man is the head of woman, and that God is the head of
the Christ. \bibverse{4} Any man who keeps his head covered, when
praying or preaching in public, dishonours him who is his head;
\bibverse{5} while any woman, who prays or preaches in public
bare-headed, dishonours him who is her head; for that is to make herself
like one of the shameless women who shave their heads. \bibverse{6}
Indeed, if a woman does not keep her head covered, she may as well cut
her hair short. But, since to cut her hair short, or shave it off, marks
her as one of the shameless women, she should keep her head covered.
\bibverse{7} A man ought not to have his head covered, for he has been
from the beginning the likeness of God and the reflection of his glory,
but woman is the reflection of man's glory. \bibverse{8} For it was not
man who was taken from woman, but woman who was taken from man.
\bibverse{9} Besides, man was not created for the sake of woman, but
woman for the sake of man. \bibverse{10} And, therefore, a woman ought
to wear on her head a symbol of her subjection, because of the presence
of the angels. \bibverse{11} Still, when in union with the Lord, woman
is not independent of man, or man of woman; \bibverse{12} for just as
woman came from man, so man comes by means of woman; and all things come
from God. \bibverse{13} judge for yourselves. Is it fitting that a woman
should pray to God in public with her head uncovered? \bibverse{14} Does
not nature herself teach us that, while for a man to wear his hair long
is degrading to him, \bibverse{15} a woman's long hair is her glory? Her
hair has been given her to serve as a covering. \bibverse{16} If,
however, anyone still thinks it right to contest the point --- well, we
have no such custom, nor have the churches of God. \bibverse{17} In
giving directions on the next subject, I cannot praise you; because your
meetings do more harm than good. \bibverse{18} To begin with, I hear you
and, to some extent, I believe it. \bibverse{19} Indeed, there must be
actual parties amongst you, for so only will the people of real worth
become known. \bibverse{20} When you meet together, as I understand, it
is not possible to eat the Lord's Supper; \bibverse{21} for, as you eat,
each of you tries to secure his own supper first, with the result that
one has too little to eat, and another has too much to drink!
\bibverse{22} Have you no houses in which you can eat and drink? Or are
you trying to show your contempt for the church of God, and to humiliate
the poor? What can I say to you? Should I praise you? In this matter I
cannot praise you. \bibverse{23} For I myself received from the Lord the
account which I have in turn given to you --- how the Lord Jesus, on the
very night of his betrayal, took some bread, \bibverse{24} and, after
saying the thanksgiving, broke it and said ``This is my own body given
on your behalf. Do this in memory of me.'' \bibverse{25} And in the same
way with the cup, after supper, saying ``This cup is the new covenant
made by my blood. Do this, whenever you drink it, in memory of me.''
\bibverse{26} For whenever you eat this bread and drink the cup, you
proclaim the Lord's death --- until he comes. \bibverse{27} Therefore,
whoever eats the bread, or drinks the Lord's cup, in an irreverent
spirit, will have to answer for an offence against the Lord's body and
blood. \bibverse{28} Let everyone look into their own heart, and only
then eat of the bread and drink from the cup. \bibverse{29} For the
person who eats and drinks brings a judgement on themselves by their
eating and drinking, when they do not discern the body. \bibverse{30}
That is why so many amongst you are weak and ill, and why some are
sleeping. \bibverse{31} But, if we judged ourselves rightly, we should
not be judged. \bibverse{32} Yet, in being judged by the Lord, we are
undergoing discipline, so that we may not have judgement passed on us
with the rest of the world. \bibverse{33} Therefore, my friends, when
you meet together to eat the Supper, wait for one another. \bibverse{34}
If anyone is hungry, they should eat at home, so that your meetings may
not bring a judgement on you. The other details I will settle when I
come.

\hypertarget{section-11}{%
\section{12}\label{section-11}}

\bibverse{1} In the next place, friends, I do not want you to be
ignorant about spiritual gifts. \bibverse{2} You know that there was a
time when you were Gentiles, going astray after idols that could not
speak, just as you happened to be led. \bibverse{3} Therefore I tell you
plainly that no one who speaks under the influence of the Spirit of God
says `JESUS IS ACCURSED,' and that no one can say `JESUS IS LORD,'
except under the influence of the Holy Spirit. \bibverse{4} Gifts
differ, but the Spirit is the same; \bibverse{5} ways of serving differ,
yet the Master is the same; \bibverse{6} results differ, yet the God who
brings about every result is in every case the same. \bibverse{7} To
each of us there is given spiritual illumination for the general good.
\bibverse{8} To one is given the power to speak with wisdom through the
Spirit; to another the power to speak with knowledge, due to the same
Spirit; \bibverse{9} to another faith by the same Spirit; to another
power to cure diseases by the one Spirit; to another supernatural
powers; \bibverse{10} to another the gift of preaching; to another the
gift of distinguishing between true and false inspiration; to another
varieties of the gift of `tongues'; to another the power to interpret
`tongues.' \bibverse{11} All these result from one and the same Spirit,
who distributes his gifts to each individually as he wills.

\bibverse{12} For just as the human body is one whole, and yet has many
parts, and all its parts, many though they are, form but one body, so it
is with the Christ; \bibverse{13} for it was by one Spirit that we were
all baptised to form one body, whether Jews or Greeks, slaves or free,
and were all imbued with one Spirit. \bibverse{14} The human body, I
repeat, consists not of one part, but of many. \bibverse{15} If the foot
says `Since I am not a hand, I do not belong to the body,' it does not
because of that cease to belong to the body. \bibverse{16} Or if the ear
says `Since I am not an eye, I do not belong to the body,' it does not
because of that cease to belong to the body. \bibverse{17} If all the
body were an eye, where would the hearing be? If it were all hearing,
where would the sense of smell be? \bibverse{18} But in fact God has
placed each individual part just where he thought fit in the body.
\bibverse{19} If, however, they all made up only one part, where would
the body be? \bibverse{20} But in fact, although it has many parts,
there is only one body. \bibverse{21} The eye cannot say to the hand `I
do not need you,' nor, again, the head to the feet `I do not need you.'
\bibverse{22} No! Those parts of the body that seem naturally the weaker
are indispensable; \bibverse{23} and those parts which we deem less
honourable we surround with special honour; and our ungraceful parts
receive a special grace which our graceful parts do not require.
\bibverse{24} Yes, God has so constructed the body --- by giving a
special honour to the part that lacks it --- \bibverse{25} As to secure
that there should be no disunion in the body, but that the parts should
show the same care for one another. \bibverse{26} If one part suffers,
all the others suffer with it, and if one part has honour done it, all
the others share its joy. \bibverse{27} Together you are the body of
Christ, and individually its parts. \bibverse{28} In the church God has
appointed, first, apostles, secondly preachers, thirdly teachers; then
he has given supernatural powers, then power to cure diseases, aptness
for helping others, capacity to govern, varieties of the gift of
`tongues.' \bibverse{29} Can everyone be an apostle? Can everyone be a
preacher? Can everyone be a teacher? Can everyone have supernatural
powers? \bibverse{30} Can everyone have power to cure diseases? Can
everyone speak in `tongues'? Can everyone interpret them? \bibverse{31}
Strive for the greater gifts. Yet I can still show you a way beyond all
comparison the best.

\hypertarget{section-12}{%
\section{13}\label{section-12}}

\bibverse{1} Though I speak in the `tongues' of people, or even of
angels, yet have not love, I have become mere echoing brass, or a
clanging cymbal! \bibverse{2} Even though I have the gift of preaching,
and fathom all hidden truths and all the depths of knowledge; even
though I have such faith as might move mountains, yet have not love, I
am nothing! \bibverse{3} Even though I dole my substance to the poor,
even though I sacrifice my body in order to boast, yet have not love, it
avails me nothing! \bibverse{4} Love is long-suffering, and kind; love
is never envious, never boastful, never conceited, never behaves
unbecomingly; \bibverse{5} love is never self-seeking, never provoked,
never reckons up her wrongs; \bibverse{6} love never rejoices at evil,
but rejoices in the triumph of truth; \bibverse{7} love bears with all
things, ever trustful, ever hopeful, ever patient. \bibverse{8} Love
never fails. But, whether it be the gift of preaching, it will be done
with; whether it be the gift of `tongues,' it will cease; whether it be
knowledge, it, too, will be done with. \bibverse{9} For our knowledge is
incomplete, and our preaching is incomplete, \bibverse{10} but, when the
perfect has come, that which is incomplete will be done with.
\bibverse{11} When I was a child, I talked as a child, I felt as a
child, I reasoned as a child; now that I am a man, I have done with
childish ways. \bibverse{12} As yet we see, in a mirror, dimly, but then
--- face to face! As yet my knowledge is incomplete, but then I will
know in full, as I have been fully known. \bibverse{13} Meanwhile faith,
hope, and love endure --- these three, but the greatest of these is
love.

\hypertarget{section-13}{%
\section{14}\label{section-13}}

\bibverse{1} Seek this love earnestly, and strive for spiritual gifts,
above all for the gift of preaching. \bibverse{2} The person who, when
speaking, uses the gift of `tongues' is speaking, not to people, but to
God, for no one understands them; yet in spirit they are speaking of
hidden truths. \bibverse{3} But those who preach are speaking to their
fellow men and women words that will build up faith, and give them
comfort and encouragement. \bibverse{4} Those who, when speaking, use
the gift of `tongues' builds up their own faith, while those who preach
build up the faith of the church. \bibverse{5} Now I want you all to
speak in `tongues,' but much more I wish that you should preach. A
preacher is worth more than one who speaks in `tongues,' unless they
interprets their words, so that the faith of the church may be built up.
\bibverse{6} This being so, friends, what good will I do you, if I come
to you and speak in `tongues,' unless my words convey some revelation,
or knowledge, or take the form of preaching or teaching? \bibverse{7}
Even with inanimate things, such as a flute or a harp, though they
produce sounds, yet unless the notes are quite distinct, how can the
tune played on the flute or the harp be recognised? \bibverse{8} If the
bugle sounds a doubtful call, who will prepare for battle? \bibverse{9}
And so with you; unless, in using the gift of `tongues,' you utter
intelligible words, how can what you say be understood? You will be
speaking to the winds! \bibverse{10} There is, for instance, a certain
number of different languages in the world, and not one of them fails to
convey meaning. \bibverse{11} If, however, I do not happen to know the
language, I will be a foreigner to those who speak it, and they will be
foreigners to me. \bibverse{12} And so with you; since your are striving
for spiritual gifts, be eager to excel in such as will build up the
faith of the church. \bibverse{13} Therefore anyone who, when speaking,
uses the gift of `tongues' should pray for ability to interpret them.
\bibverse{14} If, when praying, I use the gift of `tongues,' my spirit
indeed prays, but my mind is a blank. \bibverse{15} What, then, is my
conclusion? Simply this --- I will pray with my spirit, but with my mind
as well; I will sing with my spirit, but with my mind as well.
\bibverse{16} If you bless God with your spirit only, how can people in
the congregation who are without your gift say `Amen' to your
thanksgiving? They do not know what you are saying! \bibverse{17} Your
thanksgiving may be excellent, but the other is not helped by it.
\bibverse{18} Thank God, I use the gift of `tongues' more than any of
you. \bibverse{19} But at a meeting of the church I would rather speak
five words with my mind, and so teach others, than ten thousand words
when using the gift of `tongues.'

\bibverse{20} My friends, do not show yourselves children in
understanding. In wickedness be infants, but in understanding show
yourselves adults. \bibverse{21} It is said in the Law --- `In strange
tongues and by the lips of strangers will I speak to this people, but
even then they will not listen to me, says the Lord.' \bibverse{22}
Therefore the gift of the `tongues' is intended as a sign, not for those
who believe in Christ, but for those who do not, while the gift of
preaching is intended as a sign, not for those who do not believe in
Christ, but for those who do. \bibverse{23} So, when the whole church
meets, if all present use the gift of `tongues,' and some people who are
without the gift, or who are unbelievers, come in, will not they say
that you are mad? \bibverse{24} While, if all those present use the gift
of preaching, and an unbeliever, or someone without the gift, comes in,
they are convinced of their sinfulness by them all, they are called to
account by them all; \bibverse{25} the secrets of their heart are
revealed, and then, throwing themselves on their face, they will worship
God, and declare `God is indeed amongst you!'

\bibverse{26} What do I suggest, then, friends? Whenever you meet for
worship, each of you comes, either with a hymn, or a lesson, or a
revelation, or the gift of `tongues,' or the interpretation of them; let
everything be directed to the building up of faith. \bibverse{27} If any
of you use the gift of `tongues,' not more than two, or at the most
three, should do so --- each speaking in his turn --- and someone should
interpret them. \bibverse{28} If there is no one able to interpret what
is said, they should remain silent at the meeting of the church, and
speak to themselves and to God. \bibverse{29} Of preachers two or three
should speak, and the rest should weigh well what is said. \bibverse{30}
But, if some revelation is made to another person as he sits there, the
first speaker should stop. \bibverse{31} For you can all preach in turn,
so that all may learn some lesson and all receive encouragement.
\bibverse{32} (The spirit that moves the preachers is within the
preachers' control; \bibverse{33} for God is not a God of disorder, but
of peace.) This custom prevails in all the churches of Christ's people.
\bibverse{34} At the meetings of the church married women should remain
silent, for they are not allowed to speak in public; they should take a
subordinate place, as the Law itself directs. \bibverse{35} If they want
information on any point, they should ask their husbands about it at
home; for it is unbecoming for a married woman to speak at a meeting of
the church. \bibverse{36} What! Did God's message to the world originate
with you? Or did it find its way to none but you?

\bibverse{37} If anyone thinks that he has the gift of preaching or any
other spiritual gift, he should recognise that what I am now saying to
you is a command from the Lord. \bibverse{38} Anyone who ignores it may
be ignored. \bibverse{39} Therefore, my friends, strive for the gift of
preaching, and yet do not forbid speaking in `tongues.' \bibverse{40}
Let everything be done in a proper and orderly manner.

\hypertarget{the-apostles-teaching-as-to-the-resurrection-of-the-dead}{%
\subsection{The Apostle's Teaching as to the Resurrection of the
Dead}\label{the-apostles-teaching-as-to-the-resurrection-of-the-dead}}

\hypertarget{section-14}{%
\section{15}\label{section-14}}

\bibverse{1} Next, friends, I would like to remind you of the good news
which I told you, and which you received --- the good news on which you
have taken your stand, \bibverse{2} and by means of which you are being
saved. I would like to remind you of the words that I used in telling it
to you, since you are still holding fast to it, and since it was not in
vain that you became believers in Christ. \bibverse{3} For at the very
beginning of my teaching I gave you the account which I had myself
received --- that Christ died for our sins (as the scriptures had
foretold), \bibverse{4} that he was buried, that on the third day he was
raised (as the scriptures had foretold), \bibverse{5} and that he
appeared to Cephas, and then to the Twelve. \bibverse{6} After that, he
appeared to more than five hundred of his followers at one time, most of
whom are still alive, though some have gone to their rest. \bibverse{7}
After that, he appeared to James, and then to all the apostles.
\bibverse{8} Last of all, he appeared even to me, who am, as it were,
the abortion. \bibverse{9} For I am the meanest of the apostles, I who
am unworthy of the name of `apostle,' because I persecuted the church of
God. \bibverse{10} But it is through the love of God that I am what I
am, and the love that he showed me has not been wasted. No, I have
toiled harder than any of them, and yet it was not I, but the love of
God working with me. \bibverse{11} Whether, then, it was I or whether it
was they, this we proclaim, and this you believed. \bibverse{12} Now, if
it is proclaimed of Christ that he has been raised from the dead, how is
it that some of you say that there is no such thing as a resurrection of
the dead? \bibverse{13} But, if there is no such thing as a resurrection
of the dead, then even Christ has not been raised; \bibverse{14} and, if
Christ has not been raised, then our proclamation is without meaning,
and our faith without meaning also! \bibverse{15} Yes, and we are being
proved to have borne false testimony about God; for we testified of God
that he raised the Christ, whom he did not raise, if, indeed, the dead
do not rise! \bibverse{16} For, if the dead do not rise, then even
Christ himself has not been raised, \bibverse{17} and, if Christ has not
been raised, your faith is folly --- your sins are on you still!
\bibverse{18} Yes, and they, who have passed to their rest in union with
Christ, perished! \bibverse{19} If all that we have done has been to
place our hope in Christ for this life, then we of all people are the
most to be pitied.

\bibverse{20} But, in truth, Christ has been raised from the dead, the
first-fruits of those who are at rest. \bibverse{21} For, since through
a man there is death, so, too, through a man there is a resurrection of
the dead. \bibverse{22} For, as through union with Adam all die, so
through union with the Christ will all be made to live. \bibverse{23}
But each in their proper order --- Christ the first-fruits; afterward,
at his coming, those who belong to the Christ. \bibverse{24} Then will
come the end --- when he surrenders the kingdom to his God and Father,
having overthrown all other rule and all other authority and power.
\bibverse{25} For he must reign until God `has put all his enemies under
his feet.' \bibverse{26} The last enemy to be overthrown is death;
\bibverse{27} for God has placed all things under Christ's feet. (But,
when it is said that all things have been placed under Christ, it is
plain that God is excepted who placed everything under him.)
\bibverse{28} And, when everything has been placed under him, the Son
will place himself under God who placed everything under him, so that
God may be all in all! \bibverse{29} Again, what good will they be doing
who are baptised on behalf of the dead? If it is true that the dead do
not rise, why are people baptised on their behalf? \bibverse{30} Why,
too, do we risk our lives every hour? \bibverse{31} Daily I face death
--- I swear it, friends, by the pride in you that I feel through my
union with Christ Jesus, our Lord. \bibverse{32} If with only human
hopes I had fought in the arena at Ephesus, what should I have gained by
it? If the dead do not rise, then let us eat and drink, for tomorrow we
will die! \bibverse{33} Do not be deceived. `Good character is marred by
evil company.' \bibverse{34} Wake up to a righteous life, and cease to
sin. There are some who have no true knowledge of God. I speak in this
way to shame you. \bibverse{35} Someone, however, may ask `How do the
dead rise? And in what body will they come?' \bibverse{36} You foolish
person! The seed you yourself sow does not come to life, unless it dies!
\bibverse{37} And when you sow, you sow not the body that will be, but a
mere grain --- perhaps of wheat, or something else. \bibverse{38} God
gives it the body that he pleases --- to each seed its special body.
\bibverse{39} All forms of life are not the same; there is one for
people, another for beasts, another for birds, and another for fish.
\bibverse{40} There are heavenly bodies, and earthly bodies; but the
beauty of the heavenly bodies is not the beauty of the earthly.
\bibverse{41} There is a beauty of the sun, and a beauty of the moon,
and a beauty of the stars; for even star differs from star in beauty.
\bibverse{42} It is the same with the resurrection of the dead. Sown a
mortal body, it rises immortal; sown disfigured, it rises beautiful;
\bibverse{43} sown weak, it rises strong; sown a human body, it rises a
spiritual body. \bibverse{44} As surely as there is a human body, there
is also a spiritual body. \bibverse{45} That is what is meant by the
words --- `Adam, the first man, became a human being'; the last Adam
became a life-giving spirit. \bibverse{46} That which comes first is not
the spiritual, but the human; afterward comes the spiritual;
\bibverse{47} the first man was from the dust of the earth; the second
man from heaven. \bibverse{48} Those who are of the dust are like him
who came from the dust; and those who are of heaven are like him who
came from heaven. \bibverse{49} And as we have borne the likeness of him
who came from the dust, so let us bear the likeness of him who came from
heaven. \bibverse{50} This I say, friends --- Flesh and blood can have
no share in the kingdom of God, nor can the perishable share the
imperishable. \bibverse{51} Listen, I will tell you God's hidden
purpose! We will not all have passed to our rest, but we will all be
transformed --- in a moment, in the twinkling of an eye, \bibverse{52}
at the last trumpet-call; for the trumpet will sound, and the dead will
rise immortal, and we, also, will be transformed. \bibverse{53} For this
perishable body of ours must put on an imperishable form, and this dying
body a deathless form. \bibverse{54} And, when this dying body has put
on its deathless form, then indeed will the words of scripture come true
--- \bibverse{55} `death has been swallowed up in victory! Where, Death,
is your victory? Where, death, is your sting?' \bibverse{56} It is sin
that gives death its sting, and it is the Law that gives sin its power.
\bibverse{57} But thanks be to God, who gives us the victory, through
Jesus Christ, our Lord. \bibverse{58} Therefore, my dear friends, stand
firm, unshaken, always diligent in the Lord's work, for you know that,
in union with him, your toil is not in vain.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

\hypertarget{section-15}{%
\section{16}\label{section-15}}

\bibverse{1} With reference to the collection for Christ's people, I
want you to follow the instructions that I gave to the churches in
Galatia. \bibverse{2} On the first day of every week each of you should
put by what he can afford, so that no collections need be made after I
have come. \bibverse{3} On my arrival, I will send any persons, whom you
may authorize by letter, to carry your gift to Jerusalem; \bibverse{4}
and, if it appears to be worth while for me to go also, they will go
with me.

\bibverse{5} I will come to you as soon as I have been through Macedonia
--- for I am going through Macedonia --- \bibverse{6} And I will
probably make some stay with you or, perhaps, remain for the winter, so
that you may yourselves send me on my way, wherever I may be going.
\bibverse{7} I do not propose to pay you a visit in passing now, for I
hope to stay with you for some time, if the Lord permits. \bibverse{8} I
intend, however, staying at Ephesus until the Festival at the close of
the Harvest; \bibverse{9} for a great opening for active work has
presented itself, and there are many opponents.

\bibverse{10} If Timothy comes, take care that he has no cause for
feeling anxious while he is with you. He is doing the Master's work no
less than I am. \bibverse{11} No one, therefore, should slight him. See
him safely on his way to me, for I am expecting him with some of our
friends.

\bibverse{12} As for our friend Apollos, I have often urged him to go to
you with the others. He has, however, been very unwilling to do so as
yet; but he will go as soon as he finds a good opportunity.

\bibverse{13} Be watchful; stand firm in your faith; be brave; be
strong. \bibverse{14} Let everything you do be done in a loving spirit.

\bibverse{15} I have another request to make of you, friends. You
remember Stephanas and his household, and that they were the
first-fruits gathered in from Greece, and set themselves to serve
Christ's people. \bibverse{16} I want you, on your part, to show
deference to such people as these, as well as to every fellow labourer
and earnest worker. \bibverse{17} I am glad Stephanas and Fortunatus and
Achaicus have come, for they have made up for your absence;
\bibverse{18} they have cheered my heart, and your hearts also.
Recognise the worth of such people as these. \bibverse{19} The churches
in Roman Asia send you their greetings. Aquila and Prisca and the church
that meets at their house send you many Christian greetings.
\bibverse{20} All of the Lord's followers send you greetings. Greet one
another with a sacred kiss.

\bibverse{21} I, Paul, add this greeting in my own handwriting.
\bibverse{22} Accursed be anyone who has no love for the Lord. THE LORD
IS COMING. \bibverse{23} May the blessing of the Lord Jesus be with you.
\bibverse{24} My love to all of you who are in union with Christ Jesus.
