\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

\hypertarget{section}{%
\section{1}\label{section}}

\bibverse{1} To the church of God in Corinth, and to all Christ's people
throughout Greece, from Paul, an apostle of Christ Jesus, by the will of
God, and from Timothy, who is also a follower. \bibverse{2} May God, our
Father, and the Lord Jesus Christ bless you and give you peace.
\bibverse{3} Blessed is the God and Father of Jesus Christ our Lord, the
all-merciful Father, the God ever ready to console, \bibverse{4} who
consoles us in all our troubles, so that we may be able to console those
who are in any trouble with the consolation that we ourselves receive
from him. \bibverse{5} It is true that we have our full share of the
sufferings of the Christ, but through the Christ we have also our full
share of consolation. \bibverse{6} If we meet with trouble, it is for
the sake of your consolation and salvation; and, if we find consolation,
it is for the sake of the consolation that you will experience when you
are called to endure the sufferings that we ourselves are enduring;
\bibverse{7} and our hope for you remains unshaken. We know that, as you
are sharing our sufferings, you will also share our consolation.
\bibverse{8} We want you, friends, to know that, in the troubles which
befell us in Roman Asia, we were burdened altogether beyond our
strength, so much so that we even despaired of life. \bibverse{9}
Indeed, we had the presentiment that we must die, so that we might rely,
not on ourselves, but on God who raises the dead. \bibverse{10} And from
so imminent a death God delivered us, and will deliver us again; for in
him we have placed our hopes of future deliverance, while you, also,
help us by your prayers. \bibverse{11} And then many lips will give
thanks on our behalf for the blessing granted us in answer to many
prayers. \#\# The Apostle's Relations with his Converts

\bibverse{12} Indeed, our main ground for satisfaction is this --- Our
conscience tells us that our conduct in the world, and still more in our
relations with you, was marked by a purity of motive and a sincerity
that were inspired by God, and was based, not on worldly policy, but on
the help of God. \bibverse{13} We never write anything to you other than
what you will acknowledge to the very end --- \bibverse{14} And, indeed,
you have already partly acknowledged it about us --- that you have a
right to be proud of us, as we will be proud of you, on the day of our
Lord Jesus. \bibverse{15} With this conviction in my mind, I planned to
come to see you first, so that your pleasure might be doubled ---
\bibverse{16} To visit you both on my way to Macedonia, and to come to
you again on my return from Macedonia, and then to get you to send me on
my way into Judea. \bibverse{17} As this was my plan, where, pray, did I
show any fickleness of purpose? Or do you think that my plans are formed
on mere impulse, so that in the same breath I say `Yes' and `No'?
\bibverse{18} As God is true, the message that we brought you does not
waver between `Yes' and `No'! \bibverse{19} The Son of God, Christ
Jesus, whom we --- Silas, Timothy, and I --- proclaimed amongst you,
never wavered between `Yes' and `No.' With him it has always been `Yes.'
\bibverse{20} For, many as were the promises of God, in Christ is the
`Yes' that fulfils them. Therefore, through Christ again, let the `Amen'
rise, through us, to the glory of God. \bibverse{21} God who brings us,
with you, into close union with Christ, and who consecrated us,
\bibverse{22} also set his seal on us, and gave us his Spirit in our
hearts as a pledge of future blessings.

\bibverse{23} But, as my life will answer for it, I call God to witness
that it was to spare you that I deferred my visit to Corinth.
\bibverse{24} I do not mean that we are to dictate to you with regard to
your faith; on the contrary, we work with you for your true happiness;
indeed, it is through your faith that you are standing firm.

\hypertarget{section-1}{%
\section{2}\label{section-1}}

\bibverse{1} For my own sake, as well, I decided not to pay you another
painful visit. \bibverse{2} If it is I who cause you pain, why, who is
there to cheer me, except the person whom I am paining? \bibverse{3} So
I wrote as I did because I was afraid that if I had come, I should have
been pained by those who ought to have made me glad; for I felt sure
that it was true of you all that my joy was in every case yours also.
\bibverse{4} I wrote to you in sore trouble and distress of heart and
with many tears, not to give you pain, but to let you see how intense a
love I have for you.

\bibverse{5} Now whoever has caused the pain has not so much pained me,
as he has, to some extent --- not to be too severe --- pained every one
of you. \bibverse{6} The man to whom I refer has been sufficiently
punished by the penalty inflicted by the majority of you; \bibverse{7}
so that now you must take the opposite course, and forgive and encourage
him, or else he may be overwhelmed by the intensity of his pain.
\bibverse{8} So I entreat you to assure him of your love. \bibverse{9} I
had this further object, also, in what I wrote --- to find out whether
you might be relied on to be obedient in everything. \bibverse{10}
Anyone you forgive, I forgive them, too. Indeed, for my part, whatever I
have forgiven (if I have had to forgive anything), I have forgiven for
your sakes, in the presence of Christ, \bibverse{11} so as to prevent
Satan from taking advantage of us; for we are not ignorant of his
devices.

\bibverse{12} When I went to the district around Troas to tell the good
news of the Christ, even though there was an opening for serving the
Master, \bibverse{13} I could get no peace of mind because I failed to
find Titus, my friend; so I took leave of the people there, and went on
to Macedonia. \bibverse{14} All thanks to God, who, through our union
with the Christ, leads us in one continual triumph, and uses us to
spread the sweet perfume of the knowledge of him in every place.
\bibverse{15} For we are the fragrance of Christ ascending to God ---
both amongst those who are in the path of salvation and amongst those
who are in the path to ruin. \bibverse{16} To the latter we are a stench
which arises from death and tells of death; to the former a fragrance
which arises from life and tells of life. But who is equal to such a
task? \bibverse{17} Unlike many people, we are not in the habit of
making profit out of God's message; but in all sincerity, and bearing
God's commission, we speak before him in union with Christ.

\hypertarget{section-2}{%
\section{3}\label{section-2}}

\bibverse{1} Are we beginning to commend ourselves again? Or are we like
some who need letters of commendation to you, or from you? \bibverse{2}
You yourselves are our letter --- a letter written on our hearts, and
one which everybody can read and understand. \bibverse{3} All can see
that you are a letter from Christ delivered by us, a letter written, not
with ink, but with the Spirit of the living God, not on tablets of
stone, but on tablets of human hearts. \#\# The Ministry of the Apostles

\bibverse{4} This, then, is the confidence in regard to God that we have
gained through the Christ. \bibverse{5} I do not mean that we are fit to
form any judgement by ourselves, as if on our own authority;
\bibverse{6} our fitness comes from God, who himself made us fit to be
assistants of a new covenant, of which the substance is, not a written
Law, but a Spirit. For the written Law means death, but the Spirit gives
life.

\bibverse{7} If the system of religion which involved death, embodied in
a written Law and engraved on stones, began amid such glory, that the
Israelites were unable to gaze at the face of Moses because of its
glory, though it was but a passing glory, \bibverse{8} will not the
religion that confers the Spirit have still greater glory? \bibverse{9}
For, if there was a glory in the religion that involved condemnation,
far greater is the glory of the religion that confers righteousness!
\bibverse{10} Indeed, that which then had glory has lost its glory,
because of the glory which surpasses it. \bibverse{11} And, if that
which was to pass away was attended with glory, far more will that which
is to endure be surrounded with glory!

\bibverse{12} With such a hope as this, we speak with all plainness;
\bibverse{13} unlike Moses, who covered his face with a veil, to prevent
the Israelites from gazing at the disappearance of what was passing
away. \bibverse{14} But their minds were slow to learn. Indeed, to this
very day, at the public reading of the old covenant, the same veil
remains; only for those who are in union with Christ does it pass away.
\bibverse{15} But, even to this day, whenever Moses is read, a veil lies
on their hearts. \bibverse{16} Yet, whenever someone turns to the Lord,
the veil is removed. \bibverse{17} And the Lord is the Spirit, and,
where the Spirit of the Lord is, there is freedom. \bibverse{18} And all
of us, with faces from which the veil is lifted, seeing, as if reflected
in a mirror, the glory of the Lord, are being transformed into his
likeness, from glory to glory, as it is given by the Lord, the Spirit.

\hypertarget{section-3}{%
\section{4}\label{section-3}}

\bibverse{1} Therefore, since it is by God's mercy that we are engaged
in this ministry, we do not lose heart. \bibverse{2} No, we have
renounced the secrecy prompted by shame, refusing to adopt crafty ways,
or to tamper with God's message, and commending ourselves to everyone's
conscience, in the sight of God, by our exhibition of the truth.
\bibverse{3} And, even if the good news that we bring is veiled, it is
veiled only in the case of those who are on the path to ruin ---
\bibverse{4} people whose minds have been blinded by the God of this
age, unbelievers as they are, so that the light from the good news of
the glory of the Christ, who is the incarnation of God, should not shine
for them. \bibverse{5} (For it is not ourselves that we proclaim, but
Christ Jesus, as Lord, and ourselves as your servants for Jesus' sake.)
\bibverse{6} Indeed, the same God who said `Out of darkness light will
shine,' has shone in on our hearts, so that we should bring out into the
light the knowledge of the glory of God, seen in the face of Christ.
\bibverse{7} This treasure we have in these earthen vessels, so that its
all-prevailing power may be seen to come from God, and not to be our
own. \bibverse{8} Though hard pressed on every side, we are never hemmed
in; though perplexed, never driven to despair; \bibverse{9} though
pursued, never abandoned; though struck down, never killed!
\bibverse{10} We always bear on our bodies the marks of the death that
Jesus died, so that the life also of Jesus may be exhibited in our
bodies. \bibverse{11} Indeed, we who still live are continually being
given over to death for Jesus' sake, so that the life also of Jesus may
be exhibited in our mortal nature. \bibverse{12} And so, while death is
at work within us, life is at work within you. \bibverse{13} But, in the
same spirit of faith as that expressed in the words --- `I believed, and
therefore I spoke,' we, also believe, and therefore speak. \bibverse{14}
For we know that he who raised the Lord Jesus will raise us also with
him, and will bring us, with you, into his presence. \bibverse{15} For
all this is for your sakes, so that the loving kindness of God,
spreading from heart to heart, may cause yet more hearts to overflow
with thanksgiving, to his glory.

\bibverse{16} Therefore, as I said, we do not lose heart. No, even
though outwardly we are wasting away, yet inwardly we are being renewed
day by day. \bibverse{17} The light burden of our momentary trouble is
preparing for us, in measure transcending thought, a weight of
imperishable glory; \bibverse{18} we, all the while, gazing not on what
is seen, but on what is unseen; for what is seen is transient, but what
is unseen is imperishable.

\hypertarget{section-4}{%
\section{5}\label{section-4}}

\bibverse{1} For we know that if our tent --- that earthly body which is
now our home --- is taken down, we have a house of God's building, a
home not made by hands, imperishable, in heaven. \bibverse{2} Even while
in our present body we sigh, longing to put over it our heavenly house,
\bibverse{3} sure that, when we have put it on, we will never be found
discarnate. \bibverse{4} For we who are in this tent sigh under our
burden, unwilling to take it off, yet wishing to put our heavenly body
over it, so that all that is mortal may be absorbed in life.
\bibverse{5} And he who has prepared us for this change is God, who has
also given us his Spirit as a pledge.

\bibverse{6} Therefore we are always confident, knowing that, while our
home is in the body, we are absent from our home with the Lord.
\bibverse{7} For we guide our lives by faith, and not by what we see.
\bibverse{8} And in this confidence we would gladly leave our home in
the body, and make our home with the Lord. \bibverse{9} Therefore,
whether in our home or absent from our home, our one ambition is to
please him. \bibverse{10} For at the court of the Christ we must all
appear in our true characters, so that each may reap the results of the
life which he has lived in the body, in accordance with his actions ---
whether good or worthless. \bibverse{11} Therefore, because we know the
fear inspired by the Lord, it is true that we are trying to win people
over, but our motives are plain to God; and I hope that in your inmost
hearts they are plain to you also. \bibverse{12} We are not ``commending
ourselves'' again to you, but rather are giving you cause for pride in
us, so that you may have an answer ready for those who pride themselves
on appearances and not on character. \bibverse{13} For, if we were
``beside ourselves,'' it was in God's service! If we are not in our
senses, it is in yours! \bibverse{14} It is the love of the Christ which
compels us, when we reflect that, as one died for all, therefore all
died; \bibverse{15} and that he died for all, so that the living should
no longer live for themselves, but for him who died and rose for them.

\bibverse{16} For ourselves, then, from this time forward, we refuse to
regard anyone from the world's standpoint. Even if we once thought of
Christ from the standpoint of the world, yet now we do so no longer.
\bibverse{17} Therefore, if anyone is in union with Christ, he is a new
being! His old life has passed away; a new life has begun! \bibverse{18}
But all this is the work of God, who reconciled us to himself through
Christ, and gave us the Ministry of Reconciliation --- \bibverse{19} To
proclaim that God, in Christ, was reconciling the world to himself, not
reckoning people's offences against them, and that he had entrusted us
with the message of this reconciliation.

\bibverse{20} It is, then, on Christ's behalf that we are acting as
ambassadors, God, as it were, appealing to you through us. We implore
you on Christ's behalf --- Be reconciled to God. \bibverse{21} For our
sake God made Christ, who was innocent of sin, one with our sinfulness,
so that in him we might be made one with the righteousness of God.

\hypertarget{section-5}{%
\section{6}\label{section-5}}

\bibverse{1} Therefore, as God's fellow workers, we also appeal to you
not to receive his loving kindness in vain. \bibverse{2} For he says ---
`At the time for acceptance I listened to you, And on the day of
deliverance I helped you.' Now is the time for acceptance! Now is the
day of deliverance! \bibverse{3} Never do we put an obstacle in anyone's
way, so that no fault may be found with our ministry. \bibverse{4} No,
we are trying to commend ourselves under all circumstances, as God's
assistants should --- in many an hour of endurance, in troubles, in
hardships, in difficulties, \bibverse{5} in floggings, in imprisonments,
in riots, in toils, in sleepless nights, in fastings; \bibverse{6} by
purity, by knowledge, by patience, by kindliness, by holiness of spirit,
by unfeigned love; \bibverse{7} by the message of truth, and by the
power of God; by the weapons of righteousness in the right hand and in
the left; \bibverse{8} amid honour and disrepute, amid slander and
praise; regarded as deceivers, yet proved to be true; \bibverse{9} as
unknown, yet well-known; as at death's door, yet, see, we are living; as
chastised, yet not killed; \bibverse{10} as saddened, yet always
rejoicing; as poor, yet enriching many; as having nothing, and yet
possessing all things! \#\# The Apostle and his Converts

\bibverse{11} We have been speaking freely to you, dear friends in
Corinth; we have opened our heart; \bibverse{12} there is room there for
you, yet there is not room, in your love, for us. \bibverse{13} Can you
not in return --- I appeal to you as I should to children --- open your
hearts to us?

\bibverse{14} Do not enter into inconsistent relations with those who
reject the faith. For what partnership can there be between
righteousness and lawlessness? Or what has light to do with darkness?
\bibverse{15} What harmony can there be between Christ and Belial? Or
what can those who accept the faith have in common with those who reject
it? \bibverse{16} What agreement can you be between a temple of God and
idols? And we are a temple of the living God. That is what God meant
when he said --- `I will live amongst them, and walk amongst them; And I
will be their God, and they will be my people. \bibverse{17} Therefore
``Come out from amongst the nations, and separate yourselves from
them,'' says the Lord, ``And touch nothing impure; And I will welcome
you; \bibverse{18} and I will be a father to you, and you will be my
sons and daughters,'' says the Lord, the Ruler of all.' \# 7
\bibverse{1} With these promises, dear friends, let us purify ourselves
from everything that pollutes either body or spirit, and, in deepest
respect for God, aim at perfect holiness.

\bibverse{2} Make room for us in your hearts. In no instance have we
ever wronged, or harmed, or taken advantage of, anyone. \bibverse{3} I
am not saying this to condemn you. Indeed, I have already said that you
are in our heart, to live and die together. \bibverse{4} I have the
utmost confidence in you; I am always boasting about you. I am full of
encouragement and, in spite of all our troubles, my heart is overflowing
with happiness.

\bibverse{5} Ever since we reached Macedonia, we have had no rest in
body or mind; on every side there have been troubles --- conflicts
without, anxieties within. \bibverse{6} But God, who encourages the
downcast, has encouraged us by the arrival of Titus. \bibverse{7} And it
is not only by his arrival that we are encouraged, but also by the
encouragement which he received from you; for he tells us of your strong
affection, your penitence, and your zeal on my behalf --- so that I am
happier still. \bibverse{8} For, though I caused you sorrow by my
letter, I do not regret it. Even if I were inclined to regret it --- for
I see that my letter did cause you sorrow though only for a time ---
\bibverse{9} I am glad now; not because of the sorrow it caused you, but
because your sorrow brought you to repentance. For it was God's will
that you should feel sorrow, in order that you should not suffer loss in
any way at our hands. \bibverse{10} For, when sorrow is in accordance
with God's will, it results in a repentance leading to salvation, and
which will never be regretted. The sure result of the sorrow that the
world knows is death. \bibverse{11} For see what results that other
sorrow --- sorrow in accordance with God's will --- has had in your
case. What earnestness it produced! What explanations! What strong
feeling! What alarm! What longing! What eagerness! What readiness to
punish! You have proved yourselves altogether free from guilt in that
matter. \bibverse{12} So, then, even though I did write to you, it was
not for the sake of the wrong-doer, or of the man who was wronged, but
to make you conscious, in the sight of God, of your own earnest care for
us. And it is this that has encouraged us.

\bibverse{13} In addition to the encouragement that this gave us, we
were made far happier still by the happiness of Titus for his heart has
been cheered by you all. \bibverse{14} Although I have been boasting a
little to him about you, you did not put me to shame; but, just as every
thing we had said to you was true, so our boasting to Titus about you
has also proved to be the truth. \bibverse{15} And his affection for you
is all the greater, as he remembers the deference that you all showed
him, and recalls how you received him with anxious care. \bibverse{16} I
am glad that I can feel perfect confidence in you.

\hypertarget{the-palestine-famine-fund}{%
\subsection{The Palestine Famine Fund}\label{the-palestine-famine-fund}}

\hypertarget{section-6}{%
\section{8}\label{section-6}}

\bibverse{1} We want to remind you, friends, of the love that God has
shown to the churches in Macedonia --- \bibverse{2} How, tired though
they were by many a trouble, their overflowing happiness, and even their
deep poverty, resulted in a flood of generosity. \bibverse{3} I can bear
witness that to the full extent of their power, and even beyond their
power, spontaneously, \bibverse{4} and with many an appeal to us for
permission, they showed their love, and contributed their share towards
the fund for their fellow Christians. \bibverse{5} And that, not only in
the way we had expected; but first they gave themselves to the Lord, and
to us also, in accordance with God's will. \bibverse{6} And this led us
beg Titus, since he had started the work for you, he should also see to
the completion of this expression of your love. \bibverse{7} And,
remembering how you excel in everything --- in faith, in teaching, in
knowledge, in unfailing earnestness, and in the affection that we have
awakened in you --- I ask you to excel also in this expression of your
love.

\bibverse{8} I am not laying a command on you, but I am making use of
the earnestness shown by others to test the genuineness of your
affection. \bibverse{9} For you do not forget the loving kindness of our
Lord Jesus Christ --- how that for your sakes, although he was rich, he
became poor, so that you also might become rich through his poverty.
\bibverse{10} I am only making suggestions on this matter; for this is
the best course for you, since you were a year before others, not only
in taking action, but also in showing your readiness to do so.
\bibverse{11} And now I want you to complete the work, so that its
completion may correspond with your willing readiness --- in proportion,
of course, to your means. \bibverse{12} For, where there is willingness,
a person's gift is valued by its comparison with what they have, and not
with what they do not have. \bibverse{13} For our object is not to give
relief to others and bring distress on you. It is a matter of a fair
balance. \bibverse{14} On this occasion what you can spare will supply
what they need, at another time what they can spare may supply your
need, and so things will be equal. \bibverse{15} As scripture says ---
`Those who had much had nothing over, and those who had little did not
lack!' \bibverse{16} I thank God for inspiring Titus with the same keen
interest in your welfare that I have; \bibverse{17} for Titus has
responded to my appeals and, in his great earnestness, is starting to go
to you of his own accord. \bibverse{18} We are sending with him one of
the Lord's followers whose fame in the service of the good news has
spread through all the churches; \bibverse{19} and not only that, but he
has been elected by the churches to accompany us on our journey, in
connexion with this expression of your love, which we are personally
administering to the honour of the Lord, and to show our deep interest.
\bibverse{20} What we are specially guarding against is that any fault
should be found with us in regard to our administration of this
charitable fund; \bibverse{21} for we are trying to make arrangements
which will be right, not only in the eyes of the Lord, but also in the
eyes of people. \bibverse{22} We are also sending with them another of
our friends, whose earnestness we have many a time proved in many ways,
and whom we now find made even more earnest by his great confidence in
you. \bibverse{23} If I must say anything about Titus, he is my intimate
companion, and he shares my work for you; if it is our friends, they are
delegates of the churches, an honour to Christ. \bibverse{24} Show them,
therefore --- so that the churches may see it --- the proof of your
affection, and the ground for our boasting to them about you.

\hypertarget{section-7}{%
\section{9}\label{section-7}}

\bibverse{1} With reference, indeed, to the Fund for your fellow
Christians, it is quite superfluous for me to say anything to you.
\bibverse{2} I know, of course, your willingness to help, and I am
always boasting of it to the Macedonians. I tell them that you in Greece
have been ready for a year past; and it was really your zeal that
stimulated most of them. \bibverse{3} So my reason for sending our
friends is to prevent what we said about you from proving, in this
particular matter, an empty boast, and to enable you to be as well
prepared as I have been saying that you are. \bibverse{4} Otherwise, if
any Macedonians were to come with me, and find you unprepared, we --- to
say nothing of you --- should feel ashamed of our present confidence.
\bibverse{5} Therefore I think it necessary to beg the friends to go to
you in advance, and to complete the arrangements for the gift, which you
have already promised, so that it may be ready, as a gift, before I
come, and not look as if it were being given under pressure.

\bibverse{6} Remember the saying --- `Scanty sowing, scanty harvest;
plentiful sowing, plentiful harvest.' \bibverse{7} Let everyone give as
he has determined before hand, not grudgingly or under compulsion; for
God loves a cheerful giver. \bibverse{8} God has power to shower all
kinds of blessings on you, so that, having, under all circumstances and
on all occasions, all that you can need, you may be able to shower all
kinds of benefits on others. \bibverse{9} (As scripture says --- `He
scattered broadcast, he gave to the poor; His righteousness continues
for ever.' \bibverse{10} And he who supplies seed to the sower, and
bread for eating, will supply you with seed, and cause it to increase,
and will multiply the fruits of your righteousness). \bibverse{11} Rich
in all things yourselves, you will be able to show liberality to all,
which, with our help, will cause thanksgiving to be offered to God.
\bibverse{12} For the rendering of a public service such as this, not
only relieves the needs of your fellow Christians, but also results in
the offering to God of many a thanksgiving. \bibverse{13} Through the
evidence afforded by the service rendered, you cause people to praise
God for your fidelity to your profession of faith in the good news of
the Christ, as well as for the liberality of your contributions for them
and for all others. \bibverse{14} And they also, in their prayers for
you, express their longing to see you, because of the surpassing love of
God displayed towards you. \bibverse{15} All thanks to God for his
inestimable gift!

\hypertarget{the-apostles-claims-and-authority}{%
\subsection{The Apostle's Claims and
Authority}\label{the-apostles-claims-and-authority}}

\hypertarget{section-8}{%
\section{10}\label{section-8}}

\bibverse{1} Now, I, Paul, make a personal appeal to you by the meekness
and gentleness of the Christ --- I who, ``in your presence, am humble in
my bearing towards you, but, when absent, am bold in my language to
you'' --- \bibverse{2} I implore you not to drive me to ``show my
boldness,'' when I do come, by the confident tone which I expect to have
to adopt towards some of you, who are expecting to find us influenced in
our conduct by earthly motives. \bibverse{3} For, though we live an
earthly life, we do not wage an earthly war. \bibverse{4} The weapons
for our warfare are not earthly, but, under God, are powerful enough to
pull down strongholds. \bibverse{5} We are engaged in confuting
arguments and pulling down every barrier raised against the knowledge of
God. We are taking captive every hostile thought, to bring it into
submission to the Christ, \bibverse{6} and are fully prepared to punish
every act of rebellion, when once your submission is complete.
\bibverse{7} You look at the outward appearance of things! Let anyone,
who is confident that he belongs to Christ, reflect, for himself, again
on the fact --- that we belong to Christ no less than he does.
\bibverse{8} Even if I boast extravagantly about our authority --- which
the Lord gave us for building up your faith and not for overthrowing it
--- still I have no reason to be ashamed. \bibverse{9} I say this so
that it doesn't seem as if I am trying to overawe you by my letters.
\bibverse{10} For people say ``His letters are impressive and vigourous,
but his personal appearance is insignificant and his speaking
contemptible.'' \bibverse{11} Let such a person be assured of this ---
that our words in our letters show us to be, when absent, just what our
deeds will show us to be, when present. \bibverse{12} We have not indeed
the audacity to class or compare ourselves with some of those who
indulge in self-commendation! But, when such persons measure themselves
by themselves, and compare themselves with themselves, they show a want
of wisdom. \bibverse{13} We, however, will not give way to unlimited
boasting, but will confine ourselves to the limits of the sphere to
which God limited us, when he permitted us to come as far as Corinth.
\bibverse{14} For it is not the case, as it would be if we were not in
the habit of coming to you, that we are exceeding our bounds! Why, we
were the very first to reach you with the good news of the Christ!
\bibverse{15} Our boasting, therefore, is not unlimited, nor does it
extend to the labours of others; but our hope is that, as your faith
grows, our influence amongst you may be very greatly increased ---
though still confined to our sphere --- \bibverse{16} So that we will be
able to tell the good news in the districts beyond you, without
trespassing on the sphere assigned to others, or boasting of what has
been already done. \bibverse{17} Let anyone who boasts make their boast
of the Lord. \bibverse{18} For it is not those who commend themselves
that stand the test, but those who are commended by the Lord.

\hypertarget{section-9}{%
\section{11}\label{section-9}}

\bibverse{1} I could wish that you would tolerate a little folly in me!
But indeed you do tolerate me. \bibverse{2} I am jealous over you with
the jealousy of God. For I engaged you to one husband so that I might
present you to the Christ a pure bride. \bibverse{3} Yet I fear that it
may turn out that, just as the snake by his craftiness deceived Eve, so
your minds may have lost the loyalty and purity due from you to the
Christ. \bibverse{4} For, if some newcomer is proclaiming a Jesus other
than him whom we proclaimed, or if you are receiving a Spirit different
from the Spirit which you received, or a good news different from that
which you welcomed, then you are marvellously tolerant! \bibverse{5} I
do not regard myself as in any way inferior to the most eminent
apostles! \bibverse{6} Though I am no trained orator, yet I am not
without knowledge; indeed we made this perfectly clear to you in every
way.

\bibverse{7} Perhaps you say that I did wrong in humbling myself that
you might be exalted --- I mean because I told you God's good news
without payment. \bibverse{8} I robbed other churches by taking pay from
them, so that I might serve you! \bibverse{9} And, when I was with you
in need, I did not become a burden to any of you; for our friends, on
coming from Macedonia, supplied my needs. I kept myself, and will keep
myself from being an expense to you in any way. \bibverse{10} As surely
as I know anything of the truth of Christ, this boast, as far as I am
concerned, will not be stopped in any part of Greece. \bibverse{11} Why?
Because I do not love you? God knows that I do!

\bibverse{12} What I am doing now I will continue to do in order to cut
away the ground from under those who are wishing for some ground for
attacking me, so that as regards the thing of which they boast they may
appear in their true characters, just as we do. \bibverse{13} Such
people are false apostles, treacherous workers, disguising themselves as
apostles of Christ! \bibverse{14} And no wonder; for even Satan
disguises himself as an angel of light. \bibverse{15} It is not
surprising, therefore, if his servants also disguise themselves as
servants of righteousness. But their end will be in accordance with
their actions. \bibverse{16} I say again --- Let no one think me a fool!
Yet, if you do, at least welcome me as you would a fool, so that I, too
may indulge in a little boasting. \bibverse{17} When I speak like this,
I am not speaking as the Master would, but as a fool might, in boasting
so confidently. \bibverse{18} As so many are boasting of earthly things,
I, too, will boast. \bibverse{19} For all your cleverness, you tolerate
fools willingly enough! \bibverse{20} You tolerate a person even when
they enslave you, when they plunder you, when they get you into their
power, when they put on airs of superiority, when they strike you in the
face! \bibverse{21} I admit, to my shame, that we have been weak. But
whatever the subject on which others are not afraid to boast --- though
it is foolish to say so --- I am not afraid either! \bibverse{22} Are
they Hebrews? So am I! Are they Israelites? So am I! Are they
descendants of Abraham? So am I! \bibverse{23} Are they `Servants of
Christ'? Though it is madness to talk like this, I am more so than they!
I have had more of toil, more of imprisonment! I have been flogged times
without number. I have been often at death's door. \bibverse{24} Five
times I received at the hands of my own people forty lashes, all but
one. \bibverse{25} Three times I was beaten with rods. Once I was
stoned. Three times I was shipwrecked. I have spent a whole day and
night in the deep. \bibverse{26} My journeys have been many. I have been
through dangers from rivers, dangers from robbers, dangers from my own
people, dangers from the Gentiles, dangers in towns, dangers in the
country, dangers on the sea, dangers amongst people pretending to be
followers of the Lord. \bibverse{27} I have been through toil and
hardship. I have passed many a sleepless night; I have endured hunger
and thirst; I have often been without food; I have known cold and
nakedness. \bibverse{28} And, not to speak of other things, there is my
daily burden of anxiety about all the churches. \bibverse{29} Who is
weak without my being weak? Who is led astray without my burning with
indignation? \bibverse{30} If I must boast, I will boast of things which
show my weakness! \bibverse{31} The God and Father of the Lord Jesus ---
he who is for ever blessed --- knows that I am speaking the truth.
\bibverse{32} When I was in Damascus, the Governor under King Aretas had
the gates of that city guarded, so as to arrest me, \bibverse{33} but I
was let down in a basket through a window in the wall, and so escaped
his hands.

\hypertarget{section-10}{%
\section{12}\label{section-10}}

\bibverse{1} I must boast! It is unprofitable; but I will pass to
visions and revelations given by the Lord. \bibverse{2} I know a man in
union with Christ, who, fourteen years ago --- whether in the body or
out of the body I do not know; God knows --- was caught up (this man of
whom I am speaking) to the third heaven. \bibverse{3} And I know that
this man --- whether in the body or separated from the body I do not
know; God knows --- \bibverse{4} Was caught up into Paradise, and heard
unspeakable things of which no human being may tell. \bibverse{5} About
such a man I will boast, but about myself I will not boast except as
regards my weaknesses. \bibverse{6} Yet if I choose to boast, I will not
be a fool; for I will be speaking no more than the truth. But I refrain,
in case anyone should credit me with more than he can see in me or hear
from me, and because of the marvellous character of the revelations.
\bibverse{7} It was for this reason, and to prevent my thinking too
highly of myself, that a thorn was sent to pierce my flesh --- an
instrument of Satan to discipline me --- so that I should not think too
highly of myself. \bibverse{8} About this I three times entreated the
Lord, praying that it might leave me. \bibverse{9} But his reply has
been --- `My help is enough for you; for my strength attains its
perfection in the midst of weakness.'

Most gladly, then, will I boast all the more of my weaknesses, so that
the strength of the Christ may overshadow me. \bibverse{10} That is why
I delight in weakness, ill treatment, hardship, persecution, and
difficulties, when borne for Christ. For, when I am weak, then it is
that I am strong! \#\# Conclusion

\bibverse{11} I have been ``playing the fool!'' It is you who drove me
to it. For it is you who ought to have been commending me! Although I am
nobody, in no respect did I prove inferior to the most eminent apostles.
\bibverse{12} The marks of the true apostle were exhibited amongst you
in constant endurance, as well as by signs, by marvels, and by miracles.
\bibverse{13} In what respect, I ask, were you treated worse than the
other churches, unless it was that, for my part, I refused to become a
burden to you? Forgive me the wrong I did to you!

\bibverse{14} Remember, this is the third time that I have made every
preparation to come to see you, and I will refuse to be a burden to you;
I want, not your money, but you. It is not the duty of children to put
by for their parents, but of parents to put by for their children.
\bibverse{15} For my part, I will most gladly spend, and be spent, for
your welfare. Can it be that the more intensely I love you the less I am
to be loved? \bibverse{16} You will admit that I was not a burden to you
but you say that I was ``crafty'' and caught you ``by a trick''!
\bibverse{17} Do you assert that I took advantage of you through any of
those whom I have sent to you? \bibverse{18} I urged Titus to go, and I
sent another follower with him. Did Titus take any advantage of you?
Didn't we live in the same Spirit, and tread in the same footsteps?

\bibverse{19} Have you all this time been fancying that it is to you
that we are making our defence? No, it is in the sight of God, and in
union with Christ, that we are speaking. And all this, dear friends, is
to build up your characters; \bibverse{20} for I am afraid that perhaps,
when I come, I may find that you are not what I want you to be, and, on
the other hand, that you may find that I am what you do not want me to
be. I am afraid that I may find quarrelling, jealousy, ill feeling,
rivalry, slandering, backbiting, self-assertion, and disorder.
\bibverse{21} I am afraid that, on my next visit, my God may humble me
in regard to you, and that I may have to mourn over many who have long
been sinning, and have not repented of the impurity, immorality, and
sensuality, in which they have indulged.

\hypertarget{section-11}{%
\section{13}\label{section-11}}

\bibverse{1} For the third time I am coming to see you. By the word of
two or three witnesses each statement will be established. \bibverse{2}
I have said it, and I say it again before I come, just as if I were with
you on my second visit, though for the moment absent, I say to those who
have been long sinning, as well as to all others --- that if I come
again, I will spare no one. \bibverse{3} And that will be the proof,
which you are looking for, that the Christ speaks through me. There is
no weakness in his dealings with you. No, he shows his power amongst
you. \bibverse{4} For though his crucifixion was due to weakness, his
life is due to the power of God. And we, also, are weak in his weakness,
but with him we will live for you through the power of God. \bibverse{5}
Put yourselves to the proof, to see whether you are holding to the
faith. Test yourselves. Surely you recognise this fact about yourselves
--- that Jesus Christ is in you! Unless indeed you cannot stand the
test! \bibverse{6} But I hope that you will recognise that we can stand
the test. \bibverse{7} We pray God that you may do nothing wrong, not
that we may be seen to stand the test, but that you may do what is
right, even though we may seem not to stand the test. \bibverse{8} We
have no power at all against the truth, but we have power in the service
of the truth. \bibverse{9} We are glad when we are weak, if you are
strong. And what we pray for is that you may become perfect.
\bibverse{10} This is my reason for writing as I am now doing, while I
am away from you, so that, when I am with you, I may not act harshly in
the exercise of the authority which the Lord gave me --- and gave me for
building up and not for pulling down. \bibverse{11} And now, friends,
goodbye. Aim at perfection; take courage; agree together; live in peace.
And then God, the source of all love and peace, will be with you.
\bibverse{12} Greet one another with a sacred kiss. \bibverse{13} All
Christ's people here send you their greetings. \bibverse{14} May the
blessing of the Lord Jesus Christ, and the love of God, and the
communion with the Holy Spirit, be with you all.
