\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

\hypertarget{section}{%
\section{1}\label{section}}

\bibverse{1} To the Thessalonian church in union with God our Father and
the Lord Jesus Christ, From Paul, Silas, and Timothy. \bibverse{2} May
God, the Father, and the Lord Jesus Christ bless you and give you peace.
\#\# The Apostle and his Converts

\bibverse{3} Friends, it is our duty always to thank God about you, as
is but right, considering the wonderful growth of your faith, and
because, without exception, your love for one another is continually
increasing. \bibverse{4} So much is this the case that we ourselves
speak with pride, before the churches of God, of the patience and faith
which you have shown, in spite of all the persecutions and troubles that
you are enduring. \bibverse{5} These persecutions will vindicate the
justice of God's judgement, and will result in your being reckoned
worthy of God's kingdom, for the sake of which you are now suffering;
\bibverse{6} since God deems it just to inflict suffering on those who
are now inflicting suffering on you, \bibverse{7} and to give relief to
you who are suffering, as well as to us, at the appearing of the Lord
Jesus from heaven with his mighty angels, in flaming fire. \bibverse{8}
Then he will inflict punishment on those who refuse to know God, and on
those who turn a deaf ear to the good news of Jesus, our Lord.
\bibverse{9} They will pay the penalty of unutterable ruin --- banished
from the presence of the Lord and from the glorious manifestation of his
might, \bibverse{10} when he comes to be honoured in his people, and to
be revered in all who have learnt to believe in him (for you also
believed our testimony) --- as he will be on that day. \bibverse{11}
With this in view, our constant prayer for you is that our God may count
you worthy of the call that you have received, and by his power make
perfect your delight in all goodness and the efforts that have resulted
from your faith. \bibverse{12} Then, in the loving kindness of our God
and the Lord Jesus Christ, will the name of Jesus, our Lord, be honoured
in you, and you in him.

\hypertarget{events-that-must-precede-the-lords-coming}{%
\subsection{Events that must precede the Lord's
Coming}\label{events-that-must-precede-the-lords-coming}}

\hypertarget{section-1}{%
\section{2}\label{section-1}}

\bibverse{1} As to the coming of our Lord Jesus Christ, and our being
gathered to meet him, we beg you, friends, \bibverse{2} not lightly to
let your minds become unsettled, or disturbed by any revelation, or by
any message, or by any letter, purporting to come from us, to the effect
that the day of the Lord is come. \bibverse{3} Do not let anyone deceive
you, whatever he may do. For it will not come until after the Great
Apostasy, and the appearing of that Incarnation of Wickedness, that Lost
Soul, \bibverse{4} who so opposes himself to everyone that is spoken of
as a God or as an object of worship, and so exalts himself above them,
that he seats himself in the Temple of God, and displays himself as God!
\bibverse{5} Don't you remember how, when I was with you, I used to
speak to you of all this? \bibverse{6} And you know now what the
restraining influence is which prevents his appearing before his
appointed time. \bibverse{7} Wickedness, indeed, is already at work in
secret; but only until he who at present restrains it is removed out of
the way. \bibverse{8} Then will `Wickedness Incarnate' appear, but the
Lord Jesus will destroy him with the breath of his lips, and annihilate
him by the splendour of his coming. \bibverse{9} For at the coming of
the Lord there will be great activity on the part of Satan, in the form
of all kinds of deceptive miracles, signs, and marvels, as well as of
wicked attempts to delude --- \bibverse{10} to the ruin of those who are
on the path to destruction, because they have never received and loved
the truth to their own salvation. \bibverse{11} That is why God places
them under the influence of a delusion, to cause them to believe a lie;
\bibverse{12} so that sentence may be passed on all those who refuse to
believe the truth, but delight in wickedness. \bibverse{13} But,
friends, whom the Lord loves, it is our duty always to thank God about
you, for, from the first, God chose you for salvation through the
purifying influence of the Spirit, and your belief in the truth.
\bibverse{14} To this you were called by the good news which we brought
you, to attain to the glory of our Lord Jesus Christ. \bibverse{15}
Stand firm then, friends, and hold fast to the truths that we taught
you, whether by word or by letter. \bibverse{16} And may our Lord Jesus
Christ himself, and God our Father, who loved us and, in his loving
kindness, gave us unfailing consolation and good ground for hope,
\bibverse{17} console your hearts, and strengthen you to do and to say
all that is right.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

\hypertarget{section-2}{%
\section{3}\label{section-2}}

\bibverse{1} In conclusion, friends, pray for us --- pray that the
Lord's message may spread rapidly, and be received everywhere with
honour, as it was amongst you; \bibverse{2} and that we may be preserved
from wrong-headed and wicked people --- for it is not everyone who
believes in Christ. \bibverse{3} But the Lord will not fail you; he will
give you strength, and guard you from evil. \bibverse{4} Yes, and the
confidence that our union with the Lord enables us to place in you leads
us to believe that you are doing, and will do, what we direct you.
\bibverse{5} May the Lord bring you to the love of God, and to the
patience of the Christ. \bibverse{6} We beg you, friends, in the name of
the Lord Jesus Christ, to avoid any follower who is living an
ill-ordered life, which is not in agreement with the teaching that you
received from us. \bibverse{7} For you know well that you ought to
follow our example. When we were with you, our life was not ill-ordered,
\bibverse{8} nor did we eat anyone's bread without paying for it. Night
and day, labouring and toiling, we used to work at our trades, so as not
to be a burden on any of you. \bibverse{9} This was not because we had
not a right to receive support, but our object was to give you a pattern
for you to copy. \bibverse{10} Indeed, when we were with you, the rule
we laid down was --- `Whoever does not choose to work will not get to
eat.' \bibverse{11} We hear that there are amongst you people who are
living ill-ordered lives, and who, instead of attending to their own
business, are mere busybodies. \bibverse{12} All such people we beg, and
implore,, in the name of the Lord Jesus Christ, to attend quietly to
their business, and earn their own living. \bibverse{13} You, friends,
must not grow weary of doing what is right. \bibverse{14} If anyone
disregards what we have said in this letter, take note of them and avoid
their company, so that they may feel ashamed. \bibverse{15} Yet do not
think of them as an enemy, but caution them as you would a brother or
sister. \bibverse{16} May the Lord, from whom all peace comes, himself
give you his peace at all times and in all ways. May he be with you all.
\bibverse{17} I, Paul, add this greeting in my own handwriting. It is my
signature to every letter. This is how I write. \bibverse{18} May the
blessing of our Lord Jesus Christ be with you all.
